### 2011-12-04 paper/paper.tex
added template for paper and allowd *.tex files

\documentclass[a4paper,10pt]{llncs}

\usepackage{graphicx}
\usepackage{multicol}
\usepackage{amsbsy,amscd,amsfonts,amssymb,amstext,amsmath,latexsym,theorem}

\pagestyle{plain}
\bibliographystyle{alpha}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title{{\normalsize Seminar: Formal Specification} \\[1ex]
  Security of Multithreaded Programs by Compilation}
\author{Pascal Wittmann, Advisor: Artem Starostin}
\institute{}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

{\bf Here comes the abstract!}

\bibliography{bibliography}

\end{document}

### 2012-01-11 paper/paper.tex
added the basic structure
adjusted title
added first version of introduction

@@
 \title{{\normalsize Seminar: Formal Specification} \\[1ex]
-  Security of Multithreaded Programs by Compilation}
+  Security of Multithreaded Programs by Compilation\cite{Barthe07}}
 \author{Pascal Wittmann, Advisor: Artem Starostin}
-\institute{}
+\institute{TU Darmstadt}
 
@@
 
+\section{Introduction}
+Barthe, Rezk, Russo and Sabelfeld introduced with
+their paper \cite{Barthe07} a framework to guarantee
+noninterference in multithreaded programs at byte-code
+level. Noninterference is a security property which
+says that a program does not leak sensitive information
+to an adversary. Since more and more (mobile) devices handle
+an increasing number of tasks, multithreading is widely used
+to prevent lock ups (when e.g. establishing a 
+network-connection). But often programs that are secure when they
+are executed sequentially, leak information when they are
+composed in parallel. This leads to an attractive channel of
+information leakage.
+
+The idea of Barthe et. al. was to close this channel by compiling
+type information into the byte-code, which cause the scheduler to
+treat the secure parts in a different way. Since many mobile
+platforms already use some byte-code (e.g. Android a modification
+of the JVM byte-code).
+
+\section{Discussion of a solution}
+\section{Related Work}
+\section{Conclusion \& Outlook}
+
 \bibliography{bibliography}

### 2012-01-11 paper/paper.tex
split section 'introduction' into 'motivation' and 'introdcution to the research area'

@@
 
-\section{Introduction}
+\section{Motivation}
 Barthe, Rezk, Russo and Sabelfeld introduced with
@@
 
+\section{Introduction to research area}
 \section{Discussion of a solution}

### 2012-01-11 paper/paper.tex
added labels for sections

@@
 \section{Motivation}
+\label{sec:motivation}
 Barthe, Rezk, Russo and Sabelfeld introduced with
@@
 \section{Introduction to research area}
+\label{sec:introduction}
 \section{Discussion of a solution}
+\label{sec:discussion}
 \section{Related Work}
+\label{sec:relatedwork}
 \section{Conclusion \& Outlook}
+\label{sec:conclusion}
 

### 2012-01-11 paper/paper.tex
renamed discussion section
corrected a sentence in the motivation
added structure of section 'discussion'

@@
 type information into the byte-code, which cause the scheduler to
-treat the secure parts in a different way. Since many mobile
+treat the secure parts in a different way. The fact that many mobile
 platforms already use some byte-code (e.g. Android a modification
-of the JVM byte-code).
+of the JVM byte-code) may have been a reason for the decision to use
+type-annotated byte-code.
 
@@
 \label{sec:introduction}
-\section{Discussion of a solution}
+\section{Discussion of the paper}
 \label{sec:discussion}
+\subsection{Syntax and Semantics of multithreaded programs}
+\label{sec:syntaxsemantics}
+\subsection{Type system}
+\label{sec:typesystem}
+\subsection{Soundness}
+\label{sec:soundness}
+\subsection{Instantiation}
+\label{sec:instantiation}
+
 \section{Related Work}

### 2012-01-11 paper/paper.tex
renamed section discussion to summary
corrected spelling
added explanation of the structure of the paper

@@
 treat the secure parts in a different way. The fact that many mobile
-platforms already use some byte-code (e.g. Android a modification
+platforms already use some byte-code (e.g. Android uses a modification
 of the JVM byte-code) may have been a reason for the decision to use
@@
 \label{sec:introduction}
-\section{Discussion of the paper}
+\section{Summary of the article}
 \label{sec:discussion}
+In this summary I will follow mainly the structure of the original
+paper which is as follows. After the introduction the basic terms and
+definitions for a multithreaded programs and the scheduler are laid.
+After that the notion of security we want to achieve is presented. Along
+with this a skeleton of a type system is described, which ensures that a
+program typable in this type system is secure w.r.t the notion of security.
+The proof that this holds is sketched in the following section. In the
+last section the (by now) abstract framework is instantiated with a
+concrete example.
+
 \subsection{Syntax and Semantics of multithreaded programs}

### 2012-01-12 paper/paper.tex
first part of syntax and semantic

@@
 \label{sec:syntaxsemantics}
+A program is viewed as an abstract thing, which consists of a set of
+program points $\mathcal{P}$ with a distinguished entry ($1$) and exit
+(\texttt{exit}) point and a function that maps program points to
+instructions.
+
+These instructions are not further specified, but contain an instruction
+to start a new thread (\texttt{start \textit{pc}} where pc is the start
+instruction of the new thread).
+
+Further there is a relation $\mapsto$ that describes possible successor
+instructions. \texttt{exit} is the only instructions with no successor
+and \texttt{start \textit{pc}} may only have a single successor (the
+following program point).
+
 \subsection{Type system}

### 2012-01-13 paper/paper.tex
added citation in the explanation of the structure
paragraph about security levels

@@
 In this summary I will follow mainly the structure of the original
-paper which is as follows. After the introduction the basic terms and
-definitions for a multithreaded programs and the scheduler are laid.
-After that the notion of security we want to achieve is presented. Along
-with this a skeleton of a type system is described, which ensures that a
-program typable in this type system is secure w.r.t the notion of security.
-The proof that this holds is sketched in the following section. In the
-last section the (by now) abstract framework is instantiated with a
+paper\cite{Barthe07} which is as follows. After the introduction the basic
+terms and definitions for a multithreaded programs and the scheduler are
+laid. After that the notion of security we want to achieve is presented.
+Along with this a skeleton of a type system is described, which ensures
+that a program typable in this type system is secure w.r.t the notion of
+security. The proof that this holds is sketched in the following section.
+In the last section the (by now) abstract framework is instantiated with a
 concrete example.
@@
 
+The next thing introduced are the security levels. We assume the attacker
+'is' a level $k$. From this assumption we can reduce every set of levels
+w.l.o.g into $\{low, high\}$, where $low < high$, by mapping elements
+that are no more sensitive than $k$ to $low$ and all other elements
+including incomparable ones to $high$.
+
 \subsection{Type system}

### 2012-01-13 paper/paper.tex
corrected wording

@@
 These instructions are not further specified, but contain an instruction
-to start a new thread (\texttt{start \textit{pc}} where pc is the start
+to create a new thread (\texttt{start \textit{pc}} where pc is the start
 instruction of the new thread).

### 2012-01-13 paper/paper.tex
added commata
corrected wording

@@
 
-Further there is a relation $\mapsto$ that describes possible successor
-instructions. \texttt{exit} is the only instructions with no successor
+Further, there is a relation $\mapsto$ that describes possible successor
+instructions. \texttt{exit} is the only program point with no successor
 and \texttt{start \textit{pc}} may only have a single successor (the

### 2012-01-13 paper/paper.tex
added paragraph about security environments

@@
 
+To connect programs and security levels, a \textit{security environment}
+(se) is defined, which is used to prevent flows over implicit channels.
+A security environment is a function that maps program points to security
+levels. A program point $i$ is called high if $se(i) = high$, low if $se(i)
+= low$ and always high if all points $j$ reachable (according to $\mapsto$)
+from $i$ satisfy $se(j) = i$ and $i$ is high.
+
 \subsection{Type system}

### 2012-01-13 paper/paper.tex
formulated more precisely

@@
 = low$ and always high if all points $j$ reachable (according to $\mapsto$)
-from $i$ satisfy $se(j) = i$ and $i$ is high.
+from $i$ satisfy $se(j) = i$ and $i$ is a high program point.
+
+Now we come to the semantics part.
 

### 2012-01-13 paper/paper.tex
added paragraph about states

@@
 
-Now we come to the semantics part.
+Now we come to the semantics part. The main idea is to build a the semantics
+for multithreaded programs by combining the semantics for sequential programs
+with a scheduler.
+
+All active\footnote{a thread is active from \texttt{start \textit{pc}} until
+it reaches the \texttt{exit} point} threads are collected in a set $Thead$.
+The state of the concurrent running threads is defined as the product of
+the partial function space $(Thread \rightharpoonup LocState)$ and the set of
+global memories $GMemory$. Where $LocState$ is the internal memory of a
+thread, from there no information can leak and the $GMemory$ is the critical
+part of the system, because it is a memory shared between all active threads.
 

### 2012-01-14 paper/paper.tex
added paragraphs about 'always high' and 'hidden' threads

@@
 
+At this point a first simplification can be made. Looking a state $s$ we can
+first extract the active threads (\texttt{s.act}) by taking the domain of the first
+component. According to a security environment we can classify these threads,
+with respect to their current program point (\texttt{s.pc(tid)} where $tid \in Thread$),
+into \textit{low} threads iff the current program point is low. In \textit{high}
+threads iff the current program point is high. In \textit{always high} threads
+iff the current program point is always high and in \textit{hidden} threads, iff
+the current program point is high but not always high.
+
+The latte two are the interesting ones. If a thread is \textit{always high} it
+can not leaks any information into low, because it never only gets in touch
+with low program points. So these threads can safely be interleaved by the
+scheduler.
+
+The \textit{hidden} threads are the ones we have to care about. These contain
+obviously high information in the current program point, but have afterwards
+instructions that deal with low information. Since the attacker can watch the
+low part of the memory, chances are good that he can deduce high information
+through changing low outputs. To prevent indirect flows that are introduced by
+these hidden threads, the scheduler will be modified to treat these threads
+in a special way. This will be done by 'hiding' these threads, therefore comes
+the name of them.
+
 \subsection{Type system}

### 2012-01-14 paper/paper.tex
added paragraph about histories
added first part of scheduler definition

@@
 
+To complete our multithreaded setup we need a scheduler. The scheduler will
+operate on histories. A history is a list of pairs $(tid, l$), where $tid \in
+Thread$ and $l \in Level$. In this history all threads chosen by the scheduler
+are recorded.
+
+At this point no concrete scheduler is defined to make the framework applicable
+for a wide class of schedulers. A scheduler is in this class iff it can be
+modeled as a function $pickt: ConcState \times History \rightharpoonup Thread$
+which satisfies the following constraints:
+
 \subsection{Type system}

### 2012-01-14 paper/paper.tex
give the concurrent state a name

@@
 it reaches the \texttt{exit} point} threads are collected in a set $Thead$.
-The state of the concurrent running threads is defined as the product of
-the partial function space $(Thread \rightharpoonup LocState)$ and the set of
-global memories $GMemory$. Where $LocState$ is the internal memory of a
-thread, from there no information can leak and the $GMemory$ is the critical
-part of the system, because it is a memory shared between all active threads.
+The state of the concurrent running threads ($ConcState$) is defined as the
+product of the partial function space $(Thread \rightharpoonup LocState)$
+and the set of global memories $GMemory$. Where $LocState$ is the internal
+memory of a thread, from there no information can leak and the $GMemory$ is
+the critical part of the system, because it is a memory shared between all
+active threads.
 

### 2012-01-15 paper/paper.tex
added constraints for the scheduler (informaly)
added comment about interleaving threads
rephrased comment about interleaving threads
use package todonotes
added todo: program execution

@@
 \usepackage{amsbsy,amscd,amsfonts,amssymb,amstext,amsmath,latexsym,theorem}
+\usepackage{todonotes}
 
@@
 
+\begin{enumerate}
+\item It always picks an active thread
+\item If there is a hidden thread, always choose high or always high threads
+\item Only use low information and the low part of the history to choose a 
+      new thread
+\end{enumerate}
+
+In constraint two the interleaving of always high threads is realized.
+
+\todo{Explain how the programs are executed}
+
 \subsection{Type system}

### 2012-01-15 paper/paper.tex
added noninterference property

@@
 
+Now we can define our goal: Noninterference. We define noninterference
+in accordance to an indistinguishability relation $\sim_g$ on global memories.
+Barthe et.al. state that it is not necessary, for the purpose of the paper,
+to specify the definition of this relation. But to get a feel for this 
+relation, you can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
+\mu'|_{low}$ where $\mu|_{low}$ projects out all high elements. Based on that
+a program P is noninterfering if for all gobal memories $\mu_1, \mu_2, \mu_1'$
+and $\mu_2'$ it holds that:
+
+\begin{align*}
+\mu_1 \sim_g \mu_2 \land P,\mu_1 \Downarrow \mu_1' \land P,\mu_2 \Downarrow \mu_2' \rightarrow \mu_1' \sim_g \mu_2'
+\end{align*}
+
 \subsection{Type system}

### 2012-01-17 paper/paper.tex
removed trailing whitespaces
addded introduction to the type system
introduced sequential type system

@@
 \label{sec:typesystem}
+The type system is the core part of the framework in the effect that it
+enforces the previously defined noninterference property. Thus every
+program which is typable, is noninterfering.
+
+The type system for multithreaded programs is build up from a type system
+for sequential programs for which the following assumptions hold:
+
+\begin{enumerate}
+\item We have a partial ordered set $(LType, \leq)$ of local types
+\item and typing judgments of the form $se, i \vdash_{seq} s \Rightarrow t$, where
+$s, t \in LType$, $i \in \mathcal{P}$ and se is a securirty environment.
+\end{enumerate}
+
+This type system is extended by the following rules, to support multithreading:
+
 \subsection{Soundness}

### 2012-01-17 paper/paper.tex
use package bussproofs
added multithreading typing rules
defined a typeable program

@@
 \usepackage{todonotes}
+\usepackage{bussproofs}
 
@@
 
+\todo{Explain intuition}
 \begin{enumerate}
-\item We have a partial ordered set $(LType, \leq)$ of local types
+\item We have a partial ordered set $(LType, \leq)$ of local types, with an initial
+      type $t_{init}$
 \item and typing judgments of the form $se, i \vdash_{seq} s \Rightarrow t$, where
@@
 This type system is extended by the following rules, to support multithreading:
+
+\todo{Place proofstrees into one line with minipages}
+\begin{prooftree}
+\AxiomC{P[i] $\in$ SeqIns}
+\AxiomC{$se, i \vdash_{seq} s \Rightarrow t$}
+\BinaryInfC{$se, i \vdash s \Rightarrow t$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{P[i] = \texttt{start} pc}
+\AxiomC{$se(i) \leq se(pc)$}
+\BinaryInfC{$se, i \vdash s \Rightarrow s$}
+\end{prooftree}
+
+The first rule states, that sequential commands are treated as usual
+and the second rule ensures, that the security level of the entry point
+of the spawned thread, is lower bounded by the level of the \texttt{start}
+instruction.
+
+A program is typeable in this type system $\mathcal{S}, se \vdash P$, where a function $\mathcal{S}:
+\mathcal{P} \rightarrow LType$ that maps a local type to every program point
+and a security environment $se$, iff $\mathcal{S}$ maps every initial program point
+(including the ones of spawned threads) to $t_{init}$ and for every program point $j$ which is an
+successor of $i$ there is a type $s \in LType$, such that $s$ is lower bounded
+by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow s$ holds.
+
+\todo{Explain intuition}
 

### 2012-01-17 paper/paper.tex
use package asmth and set defintion as standard
added main theorem

@@
 \usepackage{bussproofs}
+\usepackage{amsthm}
 
+\theoremstyle{definition}
 \pagestyle{plain}
@@
 \label{sec:soundness}
+The framework is now complete, but the proof of the connection between the
+type system and the noninterference property is still outstanding. The full
+proof is not part of the paper and I will only sketch the most important
+parts.
+
+The goal is to proof the following theorem:
+
+\begin{theorem}
+If the scheduler is secure and $se, \mathcal{S} \vdash P$, then P is noninterfering
+\end{theorem}
+
 \subsection{Instantiation}

### 2012-01-17 paper/paper.tex
use package nameref
added comment about when a scheduler is secure
package asmth was in conflict with theorem
corrected spelling

@@
 \usepackage{bussproofs}
-\usepackage{amsthm}
+\usepackage{nameref}
 
-\theoremstyle{definition}
 \pagestyle{plain}
@@
 \begin{theorem}
-If the scheduler is secure and $se, \mathcal{S} \vdash P$, then P is noninterfering
+If the scheduler is secure and $se, \mathcal{S} \vdash P$, then P is noninterfering.
 \end{theorem}
+
+The scheduler is secure, if it is defined w.r.t. the conditions from section \ref{sec:syntaxsemantics}.
 

### 2012-01-17 paper/paper.tex
added intuition of the next function

@@
 
+An important hypothesis to succeed in the proof of this theorem is the existence
+of a \texttt{next} function. This \texttt{next} function should compute for every
+high program point the first subsequent program point with a low security level. With
+this function one is able to detect when a hidden thread is allowed to become visible
+again.
 \subsection{Instantiation}

### 2012-01-17 paper/paper.tex
added properties of the next function
added intuition of the first three rules

@@
 again.
+
+This intuition is capture in the following properties for the function $next: \mathcal{P}
+\rightharpoonup \mathcal{P}$:
+
+\begin{enumerate}
+\item $Dom(next) = \{i \in \mathcal{P} | se(i) = high \land \exists j \in \mathcal{P} . i \mapsto^* j \land se(j) \neq high\}$ \todo{use simple method of new paper?}
+\item $i, j \in Dom(next) \land i \mapsto j \Rightarrow next(i) = next(j)$
+\item $i \in Dom(next) \land j \not\in Dom(next) \land i \mapsto j \Rightarrow next(i) = j$
+\item $j, k \in Dom(next) \land i \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = next(k)$ \todo{is there an error?}
+\item $i, j \in Dom(next) \land k \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = k$
+\end{enumerate}
+
+Die domain of next just captures the high, but not always high program points (i.e. the ones
+that can result in a hidden thread). The second property, states that two directly successive high
+program points have the same program point in which the thread becomes visible again. The third rule
+is the counterpart of the second: If a low program point follows a high program point, this low program
+point is the result of the next function for the high program point.
+
 \subsection{Instantiation}

### 2012-01-18 paper/paper.tex
added todos
added todo
added first sentences of the instantiation and the grammar of the assambly language

@@
 
+\todo{Use names from paper}
 \begin{enumerate}
@@
 
+\todo{What is wrong with property 4 (NeP3)}
+\todo{What is wrong with property 5 (NeP4)}
+
 \subsection{Instantiation}
 \label{sec:instantiation}
+To demonstrate how the framework can be used, it was instantiated with a simple
+assembly language, which is given by the follow grammar:
+
+\todo{beautify this}
+\begin{align*}
+instr &::= binop\ op && \text{binary operation with values from stack} \\
+&|\ push\ n &&  \text{push value on the stack} \\
+&|\ load\ x && \text{push value of variable on the stack} \\
+&|\ store\ x && \text{store first element of the stack in x} \\
+&|\ goto\ j\ |\ ifeq\ j && \text{un-/conditional jump to j} \\
+&|\ start\ j && \text{create a new thread starting in j} \\
+\end{align*}
+
+where $op \in \{+,-,\times,/\}$, $n \in \mathbb{Z}$ and $x$ are variables. The
+operational semantics are standard and not explicitly necessary for the following
+instantiation, therefor they are omited.
+
+The local states are modeled as a pair of the operand stack and the program counter. The
+initial state $\lambda_{init}(pc)$ has an empty operand stack $\epsilon$ and points to
+the given initial program point.
+
+Besides this concrete language we need to define a type system to enforce noninterference
+according to \ref{sec:typesystem}. The local types are defined by a stack of security levels
+$LType = Stack(Level)$ and $t_{init}$ with the empty stack.
 

### 2012-01-18 paper/paper.tex
added figure around the extention of the sequential typing rules
forgot to add label
added one more sentence
give figure label a better name
added two examplary typing rules for the assembly language

@@
 \todo{Place proofstrees into one line with minipages}
+\begin{figure}
 \begin{prooftree}
@@
 \end{prooftree}
+\caption{Extension of the sequential typing rules}
+\label{fig:multithreaded-typing-rules}
+\end{figure}
 
@@
 according to \ref{sec:typesystem}. The local types are defined by a stack of security levels
-$LType = Stack(Level)$ and $t_{init}$ with the empty stack.
+$LType = Stack(Level)$ and $t_{init}$ with the empty stack. The typing rules defined in
+Figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
+of the assembly language. In this summary I only explain two of them, the rest follows
+in a similar manner.
+
+\todo{beautify me}
+\begin{figure}
+\begin{prooftree}
+\AxiomC{$P[i] = store\ x$}
+\AxiomC{$se(i) \sqcup k \leq \Gamma(x)$}
+\BinaryInfC{$se, i \vdash_{seq} k :: st \Rightarrow st$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$P[i] = ifeq\ j$}
+\AxiomC{$\forall j' \in reg(i), k \leq se(j')$}
+\BinaryInfC{$se, i \vdash_{seq} k :: st \Rightarrow lift_k(st)$}
+\end{prooftree}
+\caption{Excerpt of typing rules for the assembly language}
+\label{fig:transfer-rules}
+\end{figure}
 

### 2012-01-18 paper/paper.tex
explained rule for `store`

@@
 
+The security levels in $LType$ are the security levels of the operands in the local state. To
+express our security policy, we declare a function $\Gamma(x)$ that assigns to every variable
+$x$ a security level. If we now want to \texttt{store} the top of the stack into the variable
+$x$ one premise is that the security level of the program point ($se(i)$) \textit{and} the
+security level of the value on top of the operand stack is \textit{lower or equal} to the
+security level that was assigned by $Gamma$ to the variable $x$. If this is the case the head
+of $LType$ is removed.
+
 \section{Related Work}

### 2012-01-18 paper/paper.tex
corrected typo

--- 2012-01-18 paper/paper.tex
@@
 
+The security levels in $LType$ are the security levels of the operands in the local state. To
+express our security policy, we declare a function $\Gamma(x)$ that assigns to every variable
+$x$ a security level. If we now want to \texttt{store} the top of the stack into the variable
+$x$ one premise is that the security level of the program point ($se(i)$) \textit{and} the
+security level of the value on top of the operand stack is \textit{lower or equal} to the
+security level that was assigned by $\Gamma$ to the variable $x$. If this is the case the head
+of $LType$ is removed.
+
 \section{Related Work}

### 2012-01-19 paper/paper.tex
added first part of `ifeq` explanation
added paragraphs about dependence regions

@@
 
+The rule for the branching instruction can only be used if the security level of all program
+points in the control dependence region are lower bounded by the security level of the value
+that is used as the condition (i.e. the top of the operand stack). \todo{What does $lift_k$?}
+
+The next step is to construct the \texttt{next} function. To make this task easy, we introduce
+a source language which will be complied into the given assembly language. The source language
+is defined as follows:
+
+\begin{align*}
+e ::= n | x | e op e && c ::= x := e | c;c | if e then c else c | while e do c | fork(c)
+\end{align*}
+
+Now it is easy to define the control dependence region and the junction points for the source
+language and deduce the ones for the assembly language from them. The first step is to label the
+source language at the gefärdeten\todo{translate} regions with natural numbers:
+
+\begin{align*}
+c ::= [x := e]^n | c;c | [if e then c else c]^n | [while e do c]^n | [fork(c)]^n
+\end{align*}
+
+According to this labels we can define the dependence regions for the source language:
+
+\begin{definition}{sregion}
+The control dependence region for a branching command $[c]^n$ in the source language is
+defined as the labels inside the branching command, except for those ones that are inside a
+\texttt{fork} command.
+\end{definition}
+
 \section{Related Work}

### 2012-01-22 paper/paper.tex
added defintion of tregion

@@
 
+Now we can define $tregion$ according to a compilation function $\mathcal{C}$. I won't define
+the compilation function here, because it is not that essential how it exactly looks like.
+
+\begin{definition}{tregion}
+$tregion(n)$ is defined as the set of instruction labels obtained by compiling the commands
+$[c']^{n'}$ of the branching instruction $[c]^n$ of the source code with a compilation function $\mathcal{C}$.
+\end{definition}
+
 \section{Related Work}

### 2012-01-22 paper/paper.tex
added a defintion of jun

@@
 
+With this definition of $tregion$ in mind, it is easy to define the junction points.
+
+\begin{definition}{jun}
+The junction points are computed by a function $jun: \mathcal{P} \rightarrow \mathcal{P}$. It is defined
+on all junction points $[c]^n$ in the source program and its domain looks like $max \{i\ |\ i \in tregion(n)\} + 1$.
+\end{definition}
+
+Intuitively a junction point according to this definition is the point that follows on the last instruction
+that is affected by the branching instruction.
+
 \section{Related Work}

### 2012-01-22 paper/paper.tex
added paragraph about next steps to define the next function

@@
 
+This looks familiar to the \texttt{next} function we want to define.\footnote{We could define \texttt{next} for
+every instruction $i$ inside an outermost branching point $[c]^n$ as $next(i) = jun(n)$}. The part that is
+missing, is the restriction to outermost branching points whose guards involves secrets.
+
 \section{Related Work}

### 2012-01-22 paper/paper.tex
added defintion of the next function

@@
 
+To make this distinction a new type system is introduced. Type judgments have the form
+$\vdash_\alpha [c]_{\alpha'}^n : E$, where $E$ is a function that maps labels to security
+levels.\footnote{Given $E$ it is easy to define a security environment $se$. For a definition
+of $E$ see Barthe et. al. 2006} $\alpha$ denotes if $c$ is \textit{part of} a branching
+instruction that branches on secret ($\bullet$) or public ($\circ$) data and $\alpha'$ is the
+security level of the guard \textit{in} the branching instruction.
+
+We now take a look at the rules of this type system concerning the \texttt{if} instruction.
+
+\todo{use an other package to display these rules}
+\begin{figure}
+\begin{prooftree}
+\AxiomC{$\vdash e : L$}
+\AxiomC{$\vdash_\alpha c : E$}
+\AxiomC{$\vdash_\alpha c' : E$}
+\LeftLabel{L-COND:\quad}
+\TrinaryInfC{$\vdash_\alpha [if \ e \ then \ c \ else \ c']^n_\alpha : E$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\vdash e : H$}
+\AxiomC{$\vdash_\bullet c : E$}
+\AxiomC{$\vdash_\bullet c' : E$}
+\LeftLabel{H-COND:\quad}
+\TrinaryInfC{$\vdash_\bullet [if \ e \ then \ c \ else \ c']^n_\bullet : E$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\vdash e : H$ \quad $\vdash_\bullet c : E$}
+\AxiomC{$\vdash_\bullet c' : E$}
+\AxiomC{$E = lift_H(E, sregion(n))$}
+\LeftLabel{TOP-H-COND:\quad}
+\TrinaryInfC{$\vdash_\circ [if \ e \ then \ c \ else \ c']^n_\bullet : E$}
+\end{prooftree}
+
+\caption{Typing rules for \texttt{if} on source level.}
+\label{fig:typing-rules-if}
+\end{figure}
+
+The first rule L-COND from Figure \ref{fig:typing-rules-if} says that if we branch on a low
+guard, everything depends on the security level of $c$ and $c'$. The rule H-COND covers the
+case where we have a high guard, in this case the control dependence region has to be marked
+high. The rule TOP-H-COND is the interesting one. Because of the preceding rules we can not
+be part of branch with a low guard, therefore we are in the outermost high branch. The premise
+$E' = lift_H(E, sregion(n))$ reads as: For all labels in the control dependence region $E$ is
+defined as $E'(n) = H \sqcup E(n)$ and $E'(n) = E(n)$ for all other labels.
+
+This type system is powerful enough to prevent explicit and implicit flows and can therefor
+replace the type system defined previously. \todo{show ASSIGN rule?}
+
+With this type system in mind, we can now define our \texttt{next} function.
+
+\begin{definition}{next}
+For every branching point $c$ in the source program such that $\vdash_\circ [c]_\bullet^n$, we  have
+that $\forall k \in tregion(n) . next(k) = jun(n)$.
+\end{definition}
+
 \section{Related Work}

### 2012-01-23 paper/paper.tex
finisher for instantiation
translated single word

@@
 language and deduce the ones for the assembly language from them. The first step is to label the
-source language at the gefärdeten\todo{translate} regions with natural numbers:
+source language at the regions at risk with natural numbers:
 
@@
 
+The proof that this definition fulfills the properties from \nameref{sec:soundness} and all other
+proofs can be found in the full version of the paper. \todo{cite the full version}
+
+Now the instantiation of the framework is complete, except for the scheduler which was left undefined
+in the paper.
+
 \section{Related Work}

### 2012-01-23 paper/paper.tex
beautified two equatations

@@
 \begin{align*}
-e ::= n | x | e op e && c ::= x := e | c;c | if e then c else c | while e do c | fork(c)
+e ::= n\ |\ x\ |\ e\ op\ e && c\ ::=\ x\ :=\ e\ |\ c;c\ |\ if\ e\ then\ c\ else\ c\ |\ while\ e\ do\ c\ |\ fork(c)
 \end{align*}
@@
 \begin{align*}
-c ::= [x := e]^n | c;c | [if e then c else c]^n | [while e do c]^n | [fork(c)]^n
+c\ ::=\ [x\ :=\ e]^n\ |\ c;c\ |\ [if\ e\ then\ c\ else\ c]^n\ |\ [while\ e\ do\ c]^n\ |\ [fork(c)]^n
 \end{align*}

### 2012-01-23 paper/paper.tex
fixed definition of jun
use \cite for Barthe06

@@
 The junction points are computed by a function $jun: \mathcal{P} \rightarrow \mathcal{P}$. It is defined
-on all junction points $[c]^n$ in the source program and its domain looks like $max \{i\ |\ i \in tregion(n)\} + 1$.
+on all junction points $[c]^n$ in the source program and its domain is $jun(n) = max \{i\ |\ i \in tregion(n)\} + 1$.
 \end{definition}
@@
 levels.\footnote{Given $E$ it is easy to define a security environment $se$. For a definition
-of $E$ see Barthe et. al. 2006} $\alpha$ denotes if $c$ is \textit{part of} a branching
+of $E$ see \cite{Barthe06}.} $\alpha$ denotes if $c$ is \textit{part of} a branching
 instruction that branches on secret ($\bullet$) or public ($\circ$) data and $\alpha'$ is the

### 2012-01-23 paper/paper.tex
get reference in last paragraph right

@@
 
-The proof that this definition fulfills the properties from \nameref{sec:soundness} and all other
+The proof that this definition fulfills the properties from section \ref{sec:soundness} and all other
 proofs can be found in the full version of the paper. \todo{cite the full version}

### 2012-01-23 paper/paper.tex
corrected spelling of footnote

@@
 
-All active\footnote{a thread is active from \texttt{start \textit{pc}} until
-it reaches the \texttt{exit} point} threads are collected in a set $Thead$.
+All active\footnote{A thread is active from \texttt{start \textit{pc}} until
+it reaches the \texttt{exit} point.} threads are collected in a set $Thead$.
 The state of the concurrent running threads ($ConcState$) is defined as the

### 2012-01-23 paper/paper.tex
added fullstop in footnote

@@
 This looks familiar to the \texttt{next} function we want to define.\footnote{We could define \texttt{next} for
-every instruction $i$ inside an outermost branching point $[c]^n$ as $next(i) = jun(n)$}. The part that is
+every instruction $i$ inside an outermost branching point $[c]^n$ as $next(i) = jun(n)$.}. The part that is
 missing, is the restriction to outermost branching points whose guards involves secrets.

### 2012-01-23 paper/paper.tex
fixed structure
removed placeholder for abstract
cite full version of the paper correctly

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-{\bf Here comes the abstract!}
 
@@
 The proof that this definition fulfills the properties from section \ref{sec:soundness} and all other
-proofs can be found in the full version of the paper. \todo{cite the full version}
+proofs can be found in \cite{Barthe09}.
 
@@
 
-\section{Related Work}
-\label{sec:relatedwork}
-\section{Conclusion \& Outlook}
-\label{sec:conclusion}
+\section{Comparison with other aproaches/further work}
+\label{sec:furtherwork}
 

### 2012-01-23 paper/paper.tex
use package url
give the function from program points to instructions a name

@@
 \usepackage{nameref}
+\usepackage{url}
 
@@
 program points $\mathcal{P}$ with a distinguished entry ($1$) and exit
-(\texttt{exit}) point and a function that maps program points to
+(\texttt{exit}) point and a function $P$ that maps program points to
 instructions.

### 2012-01-23 paper/paper.tex
added dashes in paragraph about security levels

@@
 that are no more sensitive than $k$ to $low$ and all other elements
-including incomparable ones to $high$.
+-- including incomparable ones -- to $high$.
 

### 2012-01-23 paper/paper.tex
corrected statement about always high program points

@@
 = low$ and always high if all points $j$ reachable (according to $\mapsto$)
-from $i$ satisfy $se(j) = i$ and $i$ is a high program point.
+from $i$ satisfy $se(j) = high$ and $i$ is a high program point.
 

### 2012-01-23 paper/paper.tex
remove incorrect 'a'
corrected typo

@@
 
-Now we come to the semantics part. The main idea is to build a the semantics
+Now we come to the semantics part. The main idea is to build the semantics
 for multithreaded programs by combining the semantics for sequential programs
@@
 All active\footnote{A thread is active from \texttt{start \textit{pc}} until
-it reaches the \texttt{exit} point.} threads are collected in a set $Thead$.
+it reaches the \texttt{exit} point.} threads are collected in a set $Thread$.
 The state of the concurrent running threads ($ConcState$) is defined as the

### 2012-01-23 paper/paper.tex
corrected statement about LocState

@@
 and the set of global memories $GMemory$. Where $LocState$ is the internal
-memory of a thread, from there no information can leak and the $GMemory$ is
+memory of a thread (from there no information can leak, because everything is private/internal) and the $GMemory$ is
 the critical part of the system, because it is a memory shared between all

### 2012-01-23 paper/paper.tex
corrected grammar

--- 2012-01-23 paper/paper.tex
@@
 and the set of global memories $GMemory$. Where $LocState$ is the internal
-memory of a thread, from there no information can leak and the $GMemory$ is
+memory of a thread (from there no information can leak, because everything is private/internal) and the global memory $GMemory$ which is
 the critical part of the system, because it is a memory shared between all

### 2012-01-23 paper/paper.tex
specified which state
corrected typo

@@
 
-At this point a first simplification can be made. Looking a state $s$ we can
+At this point a first simplification can be made. Looking a state $s \in ConcState$ we can
 first extract the active threads (\texttt{s.act}) by taking the domain of the first
@@
 
-The latte two are the interesting ones. If a thread is \textit{always high} it
+The latter two are the interesting ones. If a thread is \textit{always high} it
 can not leaks any information into low, because it never only gets in touch

### 2012-01-23 paper/paper.tex
removed misplaced word
interleaved between what…

@@
 The latter two are the interesting ones. If a thread is \textit{always high} it
-can not leaks any information into low, because it never only gets in touch
-with low program points. So these threads can safely be interleaved by the
-scheduler.
+can not leaks any information into low, because it never gets in touch
+with low program points. So these threads can safely be interleaved between any
+other threads by the scheduler.
 

### 2012-01-23 paper/paper.tex
made sentence more clear
added first paragraph about execution

@@
 low part of the memory, chances are good that he can deduce high information
-through changing low outputs. To prevent indirect flows that are introduced by
+through looking at changing low outputs. To prevent indirect flows that are introduced by
 these hidden threads, the scheduler will be modified to treat these threads
@@
 \todo{Explain how the programs are executed}
+To define what noninterference means we need a notion of excution. It is assumed
+that we have a sequential execution relation $\leadsto_{seq} \subseteq SeqState \times
+SeqState$. The one step execution for the multithreaded language $\leadsto_{seq}
+\subseteq (ConcState \times History) \times (ConcState \times History)$ is defined
+by the following rules.
 

### 2012-01-23 paper/paper.tex
use package mathpartir instead of bussproof, because of bussproofs limitation to three premisses
converted bussproofs prooftrees to mathpartirs inferrules

@@
 \usepackage{bussproofs}
+\usepackage{mathpartir}
 \usepackage{nameref}
@@
 
-\todo{Place proofstrees into one line with minipages}
+\todo{Place proofstrees into one line with minipages and beautify me}
 \begin{figure}
-\begin{prooftree}
-\AxiomC{P[i] $\in$ SeqIns}
-\AxiomC{$se, i \vdash_{seq} s \Rightarrow t$}
-\BinaryInfC{$se, i \vdash s \Rightarrow t$}
-\end{prooftree}
+\begin{align*}
+\inferrule{P[i] \in SeqIns \\ se, i \vdash_{seq} s \Rightarrow t}
+{se, i \vdash s \Rightarrow t}
+\end{align*}
 
-\begin{prooftree}
-\AxiomC{P[i] = \texttt{start} pc}
-\AxiomC{$se(i) \leq se(pc)$}
-\BinaryInfC{$se, i \vdash s \Rightarrow s$}
-\end{prooftree}
+\begin{align*}
+\inferrule{P[i] = \mathtt{start} pc \\ se(i) \leq se(pc)}
+{se, i \vdash s \Rightarrow s}
+\end{align*}
 \caption{Extension of the sequential typing rules}
@@
 \begin{figure}
-\begin{prooftree}
-\AxiomC{$P[i] = store\ x$}
-\AxiomC{$se(i) \sqcup k \leq \Gamma(x)$}
-\BinaryInfC{$se, i \vdash_{seq} k :: st \Rightarrow st$}
-\end{prooftree}
+\begin{align*}
+\inferrule{P[i] = store\ x \\ se(i) \sqcup k \leq \Gamma(x)}
+{se, i \vdash_{seq} k :: st \Rightarrow st}
+\end{align*}
 
-\begin{prooftree}
-\AxiomC{$P[i] = ifeq\ j$}
-\AxiomC{$\forall j' \in reg(i), k \leq se(j')$}
-\BinaryInfC{$se, i \vdash_{seq} k :: st \Rightarrow lift_k(st)$}
-\end{prooftree}
+\begin{align*}
+\inferrule{P[i] = ifeq\ j \\ \forall j' \in reg(i), k \leq se(j')}
+{se, i \vdash_{seq} k :: st \Rightarrow lift_k(st)}
+\end{align*}
 \caption{Excerpt of typing rules for the assembly language}
@@
 \begin{figure}
-\begin{prooftree}
-\AxiomC{$\vdash e : L$}
-\AxiomC{$\vdash_\alpha c : E$}
-\AxiomC{$\vdash_\alpha c' : E$}
-\LeftLabel{L-COND:\quad}
-\TrinaryInfC{$\vdash_\alpha [if \ e \ then \ c \ else \ c']^n_\alpha : E$}
-\end{prooftree}
+\begin{align*}
+\inferrule[L-COND]{\vdash e : L \\ \vdash_\alpha c : E \\ \vdash_\alpha c' : E}
+{\vdash_\alpha [if \ e \ then \ c \ else \ c']^n_\alpha : E}
+\end{align*}
 
-\begin{prooftree}
-\AxiomC{$\vdash e : H$}
-\AxiomC{$\vdash_\bullet c : E$}
-\AxiomC{$\vdash_\bullet c' : E$}
-\LeftLabel{H-COND:\quad}
-\TrinaryInfC{$\vdash_\bullet [if \ e \ then \ c \ else \ c']^n_\bullet : E$}
-\end{prooftree}
+\begin{align*}
+\inferrule[H-COND]{\vdash e : H \\ \vdash_\bullet c : E \\ \vdash_\bullet c' : E}
+{\vdash_\bullet [if \ e \ then \ c \ else \ c']^n_\bullet : E}
+\end{align*}
 
-\begin{prooftree}
-\AxiomC{$\vdash e : H$ \quad $\vdash_\bullet c : E$}
-\AxiomC{$\vdash_\bullet c' : E$}
-\AxiomC{$E = lift_H(E, sregion(n))$}
-\LeftLabel{TOP-H-COND:\quad}
-\TrinaryInfC{$\vdash_\circ [if \ e \ then \ c \ else \ c']^n_\bullet : E$}
-\end{prooftree}
-
+\begin{align*}
+\inferrule[TOP-H-COND]{\vdash e : H \\ \vdash_\bullet c : E \\
+\vdash_\bullet c' : E \\ E = lift_H(E, sregion(n))}
+{\vdash_\circ [if \ e \ then \ c \ else \ c']^n_\bullet : E}
+\end{align*}
 \caption{Typing rules for \texttt{if} on source level.}

### 2012-01-23 paper/paper.tex
removed package bussproofs
added rules for multhreaded execution

@@
 \usepackage{todonotes}
-\usepackage{bussproofs}
 \usepackage{mathpartir}
@@
 by the following rules.
+
+\begin{figure}
+\begin{align*}
+\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in SeqIns \\
+\langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc \neq \mathtt{exit}}
+{s,h \leadsto_{conc} s[lst(ctid) := \sigma, gmem := \mu], \langle ctid, se(i)\rangle :: h}
+\end{align*}
+
+\begin{align*}
+\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in SeqIns \\
+\langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc = \mathtt{exit}}
+{s,h \leadsto_{conc} s[lst := lst \\ ctid, gmem := \mu], \langle ctid, se(i)\rangle :: h}
+\end{align*}
+
+\begin{align*}
+\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] = \mathtt{start} pc \\
+fresht_{se(i)}(s) = ntid \\ s(ctid).[pc := i + 1] = \sigma'}
+{s,h \leadsto_{conc} s.[lst(ctid) := \sigma', lst(ntid) := \lambda_{init}(pc)],
+\langle ctid, se(i)\rangle :: h}
+\end{align*}
+\caption{Multithreaded execution.}
+\label{fig:multithreaded-execution}
+\end{figure}
 

### 2012-01-23 paper/paper.tex
placed rules in miniboxes
explained first execution rule

@@
 
+The scheduler is allow to pick a new thread after every transition of $\leadsto_{seq}$. The
+intuition of the first rule is, that the scheduler picks a new thread $ctid$ which is at
+program point $i$. This program point maps to an sequential instruction. The exection of
+this instruction leads to a local state $\sigma$ and a global memory $\mu$. If these conditions
+are met and instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
+then the concurent transition can be made. In this transition the concurrent state is updated
+according to the results of the sequential execution and the thread id and security level of $i$
+is recorded in the history.
+
 Now we can define our goal: Noninterference. We define noninterference
@@
 
-\todo{Place proofstrees into one line with minipages and beautify me}
+\todo{beautify me}
 \begin{figure}
+\begin{minipage}{.5\textwidth}
 \begin{align*}
@@
 \end{align*}
-
+\end{minipage}
+\begin{minipage}{.5\textwidth}
 \begin{align*}
@@
 \end{align*}
+\end{minipage}
 \caption{Extension of the sequential typing rules}

### 2012-01-23 paper/paper.tex
explained second rule
added explanation of execution rule three

@@
 is recorded in the history.
+
+The second rule covers the case in which the thread pickted by the scheduler has only the current
+instruction before terminating (i.e. $\simga.pc = \mathtt{exit}$). In the concurrent transition
+the current thread id is removed from the concurrent state and everything else is like in the
+first rule.
+
+The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
+takes a set of thread identifiers and returns a new thread identifier at level $l$\footnote{It is
+assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$} and $\lambda_{init} : \mathcal{P}
+\rightarrow LocState$ produces an initial state with the program pointer at the given program point.
+If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
+identifier is pickt w.r.t. the security level of $i$. The program counter of the current thread
+is increased by one, to step to the next instruction. The resulting concurrent state includes the
+updated state of the current thread and the initial state for the new thread. The global memory
+is not modified and the history is extended with the current thread id and the security level of $i$.
 

### 2012-01-23 paper/paper.tex
corrected grammar
added evaluation relation

@@
 The second rule covers the case in which the thread pickted by the scheduler has only the current
-instruction before terminating (i.e. $\simga.pc = \mathtt{exit}$). In the concurrent transition
+instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
 the current thread id is removed from the concurrent state and everything else is like in the
@@
 If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
-identifier is pickt w.r.t. the security level of $i$. The program counter of the current thread
+identifier is picked w.r.t. the security level of $i$. The program counter of the current thread
 is increased by one, to step to the next instruction. The resulting concurrent state includes the
@@
 is not modified and the history is extended with the current thread id and the security level of $i$.
+
+Based on this an evaluation relation $\Downarrow_{conc} \subseteq (ConcState \times History) \times
+GMemory)$ is defined by
+
+\todo{explain intuition?}
+\begin{align*}
+s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h'. s,h \leadsto_{conc}^* s',h' \land s'.act = \emptyset \land s'.gmem = \mu
+\end{align*}
+
+$P, \mu \Downarrow_{conc} \mu'$ is a shorthand for $\langle\langle main,\lambda_{init}(1)\rangle,
+\mu\rangle, \epsilon^{hist} \Downarrow_{conc} \mu'$, where $main$ is the identity of the main thread
+and $\epsilon^{hist}$ the empty history.
 

### 2012-01-26 paper/paper.tex
corrected noninterference defintion
removed todo
beautified def of seq type system

@@
 
-\todo{Explain how the programs are executed}
 To define what noninterference means we need a notion of excution. It is assumed
@@
 \begin{align*}
-\mu_1 \sim_g \mu_2 \land P,\mu_1 \Downarrow \mu_1' \land P,\mu_2 \Downarrow \mu_2' \rightarrow \mu_1' \sim_g \mu_2'
+\mu_1 \sim_g \mu_2 \land P,\mu_1 \Downarrow_{conc} \mu_1' \land P,\mu_2 \Downarrow_{conc} \mu_2' \Rightarrow \mu_1' \sim_g \mu_2'
 \end{align*}
@@
 \begin{enumerate}
-\item We have a partial ordered set $(LType, \leq)$ of local types, with an initial
+\item We have a partial ordered set $(\text{LType}, \leq)$ of local types, with an initial
       type $t_{init}$
 \item and typing judgments of the form $se, i \vdash_{seq} s \Rightarrow t$, where
-      $s, t \in LType$, $i \in \mathcal{P}$ and se is a securirty environment.
+      $s, t \in \text{LType}$, $i \in \mathcal{P}$ and se is a securirty environment.
 \end{enumerate}

### 2012-01-26 paper/paper.tex
explained intuition of the sequential typing judgment
beautified section about typing

@@
 
-\todo{Explain intuition}
 \begin{enumerate}
 \item We have a partial ordered set $(\text{LType}, \leq)$ of local types, with an initial
-      type $t_{init}$
-\item and typing judgments of the form $se, i \vdash_{seq} s \Rightarrow t$, where
-      $s, t \in \text{LType}$, $i \in \mathcal{P}$ and se is a securirty environment.
+      type $\textsc{t}_{init}$
+\item and typing judgments of the form $se, i \vdash_{seq} \textsc{s} \Rightarrow \textsc{t}$, where
+      $\textsc{s}, \textsc{t} \in \text{LType}$, $i \in \mathcal{P}$ and se is a securirty environment.
 \end{enumerate}
+
+The intuition of the typing judgment is that if we execute the instruction at program
+point $i$ w.r.t. the security environment $se$ and the current type is $\textsc{s}$. Then the
+type after the execution is $\textsc{t}$.
 
@@
 
-\todo{beautify me}
 \begin{figure}
@@
 \begin{align*}
-\inferrule{P[i] \in SeqIns \\ se, i \vdash_{seq} s \Rightarrow t}
-{se, i \vdash s \Rightarrow t}
+\inferrule{P[i] \in \text{SeqIns} \\ se, i \vdash_{seq} \textsc{s} \Rightarrow \textsc{t}}
+{se, i \vdash \textsc{s} \Rightarrow \textsc{t}}
 \end{align*}
@@
 \begin{align*}
-\inferrule{P[i] = \mathtt{start} pc \\ se(i) \leq se(pc)}
-{se, i \vdash s \Rightarrow s}
+\inferrule{P[i] = \mathtt{start}\ pc \\ se(i) \leq se(pc)}
+{se, i \vdash \textsc{s} \Rightarrow \textsc{s}}
 \end{align*}
@@
 A program is typeable in this type system $\mathcal{S}, se \vdash P$, where a function $\mathcal{S}:
-\mathcal{P} \rightarrow LType$ that maps a local type to every program point
+\mathcal{P} \rightarrow \text{LType}$ that maps a local type to every program point
 and a security environment $se$, iff $\mathcal{S}$ maps every initial program point
-(including the ones of spawned threads) to $t_{init}$ and for every program point $j$ which is an
-successor of $i$ there is a type $s \in LType$, such that $s$ is lower bounded
-by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow s$ holds.
+(including the ones of spawned threads) to $\textsc{t}_{init}$ and for every program point $j$ which is an
+successor of $i$ there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
+by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow \textsc{s}$ holds.
 

### 2012-01-26 paper/paper.tex
beautified some more

@@
 
-At this point a first simplification can be made. Looking a state $s \in ConcState$ we can
+At this point a first simplification can be made. Looking a state $s \in \text{ConcState}$ we can
 first extract the active threads (\texttt{s.act}) by taking the domain of the first
 component. According to a security environment we can classify these threads,
-with respect to their current program point (\texttt{s.pc(tid)} where $tid \in Thread$),
+with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
 into \textit{low} threads iff the current program point is low. In \textit{high}
@@
 operate on histories. A history is a list of pairs $(tid, l$), where $tid \in
-Thread$ and $l \in Level$. In this history all threads chosen by the scheduler
+\text{Thread}$ and $l \in \text{Level}$. In this history all threads chosen by the scheduler
 are recorded.
@@
 for a wide class of schedulers. A scheduler is in this class iff it can be
-modeled as a function $pickt: ConcState \times History \rightharpoonup Thread$
+modeled as a function $pickt: \text{ConcState} \times \text{History} \rightharpoonup \text{Thread}$
 which satisfies the following constraints:
@@
 To define what noninterference means we need a notion of excution. It is assumed
-that we have a sequential execution relation $\leadsto_{seq} \subseteq SeqState \times
-SeqState$. The one step execution for the multithreaded language $\leadsto_{seq}
-\subseteq (ConcState \times History) \times (ConcState \times History)$ is defined
+that we have a sequential execution relation $\leadsto_{seq} \subseteq \text{SeqState} \times
+\text{SeqState}$. The one step execution for the multithreaded language $\leadsto_{seq}
+\subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
 by the following rules.
@@
 \begin{align*}
-\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in SeqIns \\
+\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in \text{SeqIns} \\
 \langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc \neq \mathtt{exit}}
@@
 \begin{align*}
-\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in SeqIns \\
+\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] \in \text{SeqIns} \\
 \langle s(ctid), s.gmem\rangle \leadsto_{seq} \sigma,\mu \\ \sigma.pc = \mathtt{exit}}
@@
 The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
-takes a set of thread identifiers and returns a new thread identifier at level $l$\footnote{It is
-assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$} and $\lambda_{init} : \mathcal{P}
-\rightarrow LocState$ produces an initial state with the program pointer at the given program point.
+\footnote{It is assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$}
+takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}
+\rightarrow \text{LocState}$ produces an initial state with the program pointer at the given program point.
 If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
@@
 
-Based on this an evaluation relation $\Downarrow_{conc} \subseteq (ConcState \times History) \times
-GMemory)$ is defined by
+Based on this an evaluation relation $\Downarrow_{conc} \subseteq (\text{ConcState} \times \text{History}) \times
+\text{GMemory})$ is defined by
 
@@
 
-\todo{beautify this}
 \begin{align*}
-instr &::= binop\ op && \text{binary operation with values from stack} \\
-&|\ push\ n &&  \text{push value on the stack} \\
-&|\ load\ x && \text{push value of variable on the stack} \\
-&|\ store\ x && \text{store first element of the stack in x} \\
-&|\ goto\ j\ |\ ifeq\ j && \text{un-/conditional jump to j} \\
-&|\ start\ j && \text{create a new thread starting in j} \\
+instr &::= \text{binop}\ op && \text{binary operation with values from stack} \\
+&|\ \text{push}\ n &&  \text{push value on the stack} \\
+&|\ \text{load}\ x && \text{push value of variable on the stack} \\
+&|\ \text{store}\ x && \text{store first element of the stack in $x$} \\
+&|\ \text{goto}\ j\ |\ \text{ifeq}\ j && \text{un-/conditional jump to $j$} \\
+&|\ \text{start}\ j && \text{create a new thread starting in $j$} \\
 \end{align*}
@@
 Besides this concrete language we need to define a type system to enforce noninterference
-according to \ref{sec:typesystem}. The local types are defined by a stack of security levels
-$LType = Stack(Level)$ and $t_{init}$ with the empty stack. The typing rules defined in
-Figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
+according to section \ref{sec:typesystem}. The local types are defined by a stack of security levels
+$\text{LType} = \text{Stack}(\text{Level})$ and $\textsc{t}_{init}$ with the empty stack. The typing rules defined in
+figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
 of the assembly language. In this summary I only explain two of them, the rest follows
@@
 
-\todo{beautify me}
 \begin{figure}
+\begin{minipage}{.5\textwidth}
 \begin{align*}
-\inferrule{P[i] = store\ x \\ se(i) \sqcup k \leq \Gamma(x)}
+\inferrule{P[i] = \text{store}\ x \\ se(i) \sqcup k \leq \Gamma(x)}
 {se, i \vdash_{seq} k :: st \Rightarrow st}
 \end{align*}
-
+\end{minipage}
+\begin{minipage}{.5\textwidth}
 \begin{align*}
-\inferrule{P[i] = ifeq\ j \\ \forall j' \in reg(i), k \leq se(j')}
+\inferrule{P[i] = \text{ifeq}\ j \\ \forall j' \in reg(i), k \leq se(j')}
 {se, i \vdash_{seq} k :: st \Rightarrow lift_k(st)}
 \end{align*}
-\caption{Excerpt of typing rules for the assembly language}
+\end{minipage}
+\caption{Excerpt of typing rules for the assembly language.}
 \label{fig:transfer-rules}
@@
 \begin{align*}
-e ::= n\ |\ x\ |\ e\ op\ e && c\ ::=\ x\ :=\ e\ |\ c;c\ |\ if\ e\ then\ c\ else\ c\ |\ while\ e\ do\ c\ |\ fork(c)
+e ::= n\ |\ x\ |\ e\ op\ e && c ::= x\ :=\ e\ |\ c;c\ |\ \text{if}\ e\ \text{then}\ c\ \text{else}\ c\ |\ \text{while}\ e\ \text{do}\ c\ |\ \text{fork}(c)
 \end{align*}
@@
 \begin{align*}
-c\ ::=\ [x\ :=\ e]^n\ |\ c;c\ |\ [if\ e\ then\ c\ else\ c]^n\ |\ [while\ e\ do\ c]^n\ |\ [fork(c)]^n
+c ::= [x\ :=\ e]^n\ |\ c;c\ |\ [\text{if}\ e\ \text{then}\ c\ \text{else}\ c]^n\ |\ [\text{while}\ e\ \text{do}\ c]^n\ |\ [\text{fork}(c)]^n
 \end{align*}
@@
 \begin{align*}
-\inferrule[L-COND]{\vdash e : L \\ \vdash_\alpha c : E \\ \vdash_\alpha c' : E}
-{\vdash_\alpha [if \ e \ then \ c \ else \ c']^n_\alpha : E}
+\inferrule[L-Cond]{\vdash e : L \\ \vdash_\alpha c : E \\ \vdash_\alpha c' : E}
+{\vdash_\alpha [\text{if} \ e \ \text{then} \ c \ \text{else} \ c']^n_\alpha : E}
 \end{align*}
@@
 \begin{align*}
-\inferrule[H-COND]{\vdash e : H \\ \vdash_\bullet c : E \\ \vdash_\bullet c' : E}
-{\vdash_\bullet [if \ e \ then \ c \ else \ c']^n_\bullet : E}
+\inferrule[H-Cond]{\vdash e : H \\ \vdash_\bullet c : E \\ \vdash_\bullet c' : E}
+{\vdash_\bullet [\text{if} \ e \ \text{then} \ c \ \text{else} \ c']^n_\bullet : E}
 \end{align*}
@@
 \begin{align*}
-\inferrule[TOP-H-COND]{\vdash e : H \\ \vdash_\bullet c : E \\
+\inferrule[Top-H-Cond]{\vdash e : H \\ \vdash_\bullet c : E \\
 \vdash_\bullet c' : E \\ E = lift_H(E, sregion(n))}
-{\vdash_\circ [if \ e \ then \ c \ else \ c']^n_\bullet : E}
+{\vdash_\circ [\text{if} \ e \ \text{then} \ c \ \text{else} \ c']^n_\bullet : E}
 \end{align*}
@@
 
-The first rule L-COND from Figure \ref{fig:typing-rules-if} says that if we branch on a low
-guard, everything depends on the security level of $c$ and $c'$. The rule H-COND covers the
+The first rule \textsc{L-Cond} from Figure \ref{fig:typing-rules-if} says that if we branch on a low
+guard, everything depends on the security level of $c$ and $c'$. The rule \textsc{H-Cond} covers the
 case where we have a high guard, in this case the control dependence region has to be marked
-high. The rule TOP-H-COND is the interesting one. Because of the preceding rules we can not
+high. The rule \textsc{Top-H-Cond} is the interesting one. Because of the preceding rules we can not
 be part of branch with a low guard, therefore we are in the outermost high branch. The premise

### 2012-01-26 paper/paper.tex
corrected typo

@@
 This looks familiar to the \texttt{next} function we want to define.\footnote{We could define \texttt{next} for
-every instruction $i$ inside an outermost branching point $[c]^n$ as $next(i) = jun(n)$.}. The part that is
+every instruction $i$ inside an outermost branching point $[c]^n$ as $next(i) = jun(n)$.} The part that is
 missing, is the restriction to outermost branching points whose guards involves secrets.

### 2012-01-26 paper/paper.tex
removed todo about mathpartir
used minipaged in advanced typing rules to save space
exlained intution of eval beautified eval removed todo
corrected spelling

@@
 
-\todo{explain intuition?}
 \begin{align*}
-s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h'. s,h \leadsto_{conc}^* s',h' \land s'.act = \emptyset \land s'.gmem = \mu
+s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h'.\ (s,h \leadsto_{conc}^* s',h') \land s'.act = \emptyset \land s'.gmem = \mu
 \end{align*}
+
+So the state $s$ evaluates according to a history $h$ to a final global memory $\mu$ iff there is a
+sequence of concurrent executions that terminates (i.e. there are no active threads) and has the
+global memory $\mu$.
 
@@
 
-A program is typeable in this type system $\mathcal{S}, se \vdash P$, where a function $\mathcal{S}:
-\mathcal{P} \rightarrow \text{LType}$ that maps a local type to every program point
-and a security environment $se$, iff $\mathcal{S}$ maps every initial program point
+A program is typeable in this type system (written $\mathcal{S}, se \vdash P$),
+where $\mathcal{S}$ is a function $\mathcal{S}: \mathcal{P} \rightarrow \text{LType}$
+that maps a local type to every program point
+and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point
 (including the ones of spawned threads) to $\textsc{t}_{init}$ and for every program point $j$ which is an
@@
 
-\todo{use an other package to display these rules}
 \begin{figure}
+\begin{minipage}{.5\textwidth}
 \begin{align*}
@@
 \end{align*}
+\end{minipage}
 

### 2012-01-26 paper/paper.tex
removed todo
added remark about \sqcup
explained lift_k

@@
 security level that was assigned by $\Gamma$ to the variable $x$. If this is the case the head
-of $LType$ is removed.
+of $LType$ is removed. The $\sqcup$ operator is like a logical 'or', it returns the higher of
+the given security values.
 
@@
 points in the control dependence region are lower bounded by the security level of the value
-that is used as the condition (i.e. the top of the operand stack). \todo{What does $lift_k$?}
+that is used as the condition (i.e. the top of the operand stack). The function $lift_k :
+\text{Level} \rightarrow \text{Stack}(\text{Level}) \rightarrow \text{Stack}(\text{Level})$
+extends $\sqcup$ to stacks of security levels. This ensures, that every program point in the control
+dependence region has the same security level as the branching point.
 
@@
 This type system is powerful enough to prevent explicit and implicit flows and can therefor
-replace the type system defined previously. \todo{show ASSIGN rule?}
+replace the type system defined previously.
 

### 2012-01-26 paper/paper.tex
added fullstop in footnote

@@
 The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
-\footnote{It is assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$}
+\footnote{It is assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$.}
 takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}

### 2012-01-26 paper/paper.tex
added fullstop in figure
added explicit space in definition of next

@@
 \end{minipage}
-\caption{Extension of the sequential typing rules}
+\caption{Extension of the sequential typing rules.}
 \label{fig:multithreaded-typing-rules}
@@
 For every branching point $c$ in the source program such that $\vdash_\circ [c]_\bullet^n$, we  have
-that $\forall k \in tregion(n) . next(k) = jun(n)$.
+that $\forall k \in tregion(n) .\ next(k) = jun(n)$.
 \end{definition}

### 2012-01-27 paper/paper.tex
added note about reflexsive, transitive closure of \leadsto_{conc}
be more explicit

@@
 The next thing introduced are the security levels. We assume the attacker
-'is' a level $k$. From this assumption we can reduce every set of levels
+'is' a level $k$. From this assumption we can reduce every set of security levels
 w.l.o.g into $\{low, high\}$, where $low < high$, by mapping elements
@@
 sequence of concurrent executions that terminates (i.e. there are no active threads) and has the
-global memory $\mu$.
+global memory $\mu$. $\leadsto_{conc}^*$ is the reflexive and transitive closure of of $\leadsto_{conc}$.
 

### 2012-01-27 paper/paper.tex
fixed introduction section title
added first paragraph of introduction

@@
 
-\section{Introduction to research area}
+\section{Introduction to the research area}
 \label{sec:introduction}
+The challenge to protect the confidentiality of information processed by
+a computer lead thirty years ago (c.f. \cite{Zdancewic04}) to the research
+on the security information-flows, because of the shortcomings of the standard methods.
+These standard methods are: Access control and encryption. Where
+access control ensures that the data is only accessed by an authorized entity,
+and encryption ensures that data can be securely transmitted over an insecure channel.
+But all these methods cover the release of information and not the further
+propagation (i.e. how the information is used after having access to it or decrypting
+it). This is the place where information-flow security comes into play.
+
 \section{Summary of the article}

### 2012-01-28 paper/paper.tex
added categories of side channels

@@
 it). This is the place where information-flow security comes into play.
+
+The idea is to 'track and regulate' \cite{Zdancewic04} where information flows to
+prevent the leak of secret information. The ways on which the information flows
+through the system are called channels. Channels that are not indented to transport
+information are called covered channels\cite{Sabelfeld03}. These covered channels can
+be classified into the following categories (in the following examples $h$ is a secret
+variable and $l$ a variable that is plublic):
+
+\begin{itemize}
+\item \textit{Explicit and implicit flows}: In an explicit flow the information is leaked
+      directly into some public variable (e.g. $l := h$), while in implicit flows the information
+      is leaked through the control structure of the program (e.g. $\text{if}\ h = 1\ \text{then}\ l := 1\ \text{else}\ l := 0$).
+\item \textit{Probabilistic channels}: If an attacker is able to run a computation multiple times
+      he might be able to obtain information by looking at the probability distribution of the
+      public outputs.
+\item \textit{Power channels}: If the attacker has physical access to the computer or at least figure out
+      the power consumption, he might be able to obtain data through the changing power consumption.
+\item \textit{Resource exhausting channels}: Information also may be leaked through the exhaustion
+      of finite (mostly physical) resources (e.g. a buffer overflow).
+\item \textit{Termination channels}: The attacker can obtain information through the termination
+      or non-termination of the computation or program.
+\item \textit{Timing channels:} Information can be obtained through the time at which this action
+      occurs. On the one hand \textit{external} timing channels cover the action like the termination
+      of the program, where the attacker obtains information from the total execution time of the program.
+      On the other hand \textit{internal} timing channels are occur in multithreaded programs through the timing difference
+      between threads. For example the program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
+      executed with most schedulers will leak the information of $h$ into $l$ through the timing difference
+      if $h = 1$.
+\end{itemize}
 

### 2012-01-28 paper/paper.tex
fixed typo in footnote

@@
 The third rule introduces the dynamic creation of new threads. In this rule the function $fresh_{l}$
-\footnote{It is assumed that $fresh_l \neq fresh_{l'} Leftrightarrow l \neq l'$.}
+\footnote{It is assumed that $fresh_l(tid) \neq fresh_{l'}(tid) \Leftrightarrow l \neq l'$.}
 takes a set of thread identifiers and returns a new thread identifier at level $l$ and $\lambda_{init} : \mathcal{P}

### 2012-01-28 paper/paper.tex
fixed the example for internal timing channels

@@
       On the other hand \textit{internal} timing channels are occur in multithreaded programs through the timing difference
-      between threads. For example the program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
+      between threads. For example the two threaded (modeled with $||$) program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
       executed with most schedulers will leak the information of $h$ into $l$ through the timing difference

### 2012-01-28 paper/paper.tex
removed trailing whitespace
fixed citation

@@
 through the system are called channels. Channels that are not indented to transport
-information are called covered channels\cite{Sabelfeld03}. These covered channels can
-be classified into the following categories (in the following examples $h$ is a secret
+information are called covered channels. These covered channels can
+be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
 variable and $l$ a variable that is plublic):

### 2012-01-28 paper/paper.tex
added paragraph about noninterference and declassification

@@
 \end{itemize}
+
+The notion of confidentiality that is used in most work in the area of information-flow security is \textit{noninterference}.
+This policy is defined in various ways but the essence is, that it requires 'that secret information
+[do] not affect [the] publicly observable behavior of a system'\cite{Zdancewic04}. For most real world
+applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{For a
+password checker its necessary to reveal whether the user input was correct or not.} Approaches that allow
+controlled release of information are called declassification. The kinds of declassification are \textit{what}
+information is released, \textit{who} released this information, \textit{where} in the system the information
+is released and \textit{when} the information is released (c.f. \cite{Sabelfeld05}). For these approaches it is necessary that an
+active attacker can only know as much as a passive attacker.
+
+The paper resumed in the following concentrates on internal timing channels with a noninterference policy.
 

### 2012-01-28 paper/paper.tex
added paragraph about static and dynamic analysis

@@
 active attacker can only know as much as a passive attacker.
+
+Mechanisms for controlling the information flow can be implemented either dynamic or static. The dynamic approach\footnote{
+The taint mode of the programming language Perl uses this mechanism.}
+labels the information with security labels and propagates these labels wherever the information is used. The static
+approach\footnote{This approach is implemented in Jif for Java and Flow Caml for Caml.} analyses the program code and is
+therefor far more promising, because with this approach it is possible to check all evaluation paths.
 

### 2012-01-28 paper/paper.tex
rephrased whole motivation section

@@
 \label{sec:motivation}
-Barthe, Rezk, Russo and Sabelfeld introduced with
-their paper \cite{Barthe07} a framework to guarantee
-noninterference in multithreaded programs at byte-code
-level. Noninterference is a security property which
-says that a program does not leak sensitive information
-to an adversary. Since more and more (mobile) devices handle
-an increasing number of tasks, multithreading is widely used
-to prevent lock ups (when e.g. establishing a
-network-connection). But often programs that are secure when they
-are executed sequentially, leak information when they are
-composed in parallel. This leads to an attractive channel of
-information leakage.
+Protecting the confidentiality of information that is processed by
+a modern computer is challenging and important, since more and more sensitive
+information is fed to them. Access control and encryption are not enough to
+ensure this confidentiality, because the usage of the information after the
+access or decryption is not restricted. It is necessary to control the
+\textit{information flow} to protect the sensitive information.
+On modern mobile devices sensitive information is processed and programs
+are executed multithreaded (e.g. to prevent lock ups when establishing
+a network connection). Through the timing difference of the scheduled threads it
+is likely that sensitive information is leaked and an attack may obtain all
+sensitive data.
 
-The idea of Barthe et. al. was to close this channel by compiling
-type information into the byte-code, which cause the scheduler to
-treat the secure parts in a different way. The fact that many mobile
-platforms already use some byte-code (e.g. Android uses a modification
-of the JVM byte-code) may have been a reason for the decision to use
-type-annotated byte-code.
+The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
+this leak (more formally called \textit{covert channel}) by annotating the byte-code
+of the program (most mobile devices use some sort of byte-code e.g. Android) with security
+labels according to the processed information, which causes the scheduler to hide
+threads that leak information from the attacker.
 

### 2012-01-28 paper/paper.tex
added a remark in introduction and shortened a footnote

@@
 This policy is defined in various ways but the essence is, that it requires 'that secret information
-[do] not affect [the] publicly observable behavior of a system'\cite{Zdancewic04}. For most real world
-applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{For a
-password checker its necessary to reveal whether the user input was correct or not.} Approaches that allow
+[do] not affect [the] publicly observable behavior of a system'\footnote{Even wen the attacker has full access to the
+source code of the program.}\cite{Zdancewic04}. For most real world
+applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{A
+password checker needs to reveal whether the user input was correct or not.} Approaches that allow
 controlled release of information are called declassification. The kinds of declassification are \textit{what}

### 2012-01-28 paper/paper.tex
s/undefined/unspecified
fixed wording

@@
       On the other hand \textit{internal} timing channels are occur in multithreaded programs through the timing difference
-      between threads. For example the two threaded (modeled with $||$) program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
+      between threads. For example the two threaded (denoted with $||$) program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
       executed with most schedulers will leak the information of $h$ into $l$ through the timing difference
@@
 
-Now the instantiation of the framework is complete, except for the scheduler which was left undefined
+Now the instantiation of the framework is complete, except for the scheduler which was left unspecified
 in the paper.

### 2012-01-28 paper/paper.tex
added one more property about the attacker
colon instead of fullstop in formular
beautified definitions

@@
 that are no more sensitive than $k$ to $low$ and all other elements
--- including incomparable ones -- to $high$.
+-- including incomparable ones -- to $high$. It is also assumed, that access
+control works correctly (i.e. the attacker can not access $high$ elements
+directly).
 
@@
 \begin{align*}
-s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h'.\ (s,h \leadsto_{conc}^* s',h') \land s'.act = \emptyset \land s'.gmem = \mu
+s, h \Downarrow_{conc} \mu \Leftrightarrow \exists s', h':\ (s,h \leadsto_{conc}^* s',h') \land s'.act = \emptyset \land s'.gmem = \mu
 \end{align*}
@@
 
-\begin{definition}{sregion}
-The control dependence region for a branching command $[c]^n$ in the source language is
+\begin{definition}
+The control dependence region $sregion(n)$ for a branching command $[c]^n$ in the source language is
 defined as the labels inside the branching command, except for those ones that are inside a
@@
 
-\begin{definition}{tregion}
+\begin{definition}
 $tregion(n)$ is defined as the set of instruction labels obtained by compiling the commands
@@
 
-\begin{definition}{jun}
+\begin{definition}
 The junction points are computed by a function $jun: \mathcal{P} \rightarrow \mathcal{P}$. It is defined
@@
 
-\begin{definition}{next}
-For every branching point $c$ in the source program such that $\vdash_\circ [c]_\bullet^n$, we  have
-that $\forall k \in tregion(n) .\ next(k) = jun(n)$.
+\begin{definition}
+For every branching point $c$ in the source program such that $\vdash_\circ [c]_\bullet^n$, the next function
+is defined as $\forall k \in tregion(n) .\ next(k) = jun(n)$.
 \end{definition}

### 2012-01-28 paper/paper.tex
fixed spelling error in section title
first paragraph of section comparison
added \section's of the two approaches I will compare to

@@
 
-\section{Comparison with other aproaches/further work}
+\section{Comparison with other approaches/further work}
 \label{sec:furtherwork}
+In this section I will first give a short overview of two other approaches and then
+I will compare them with the approach presented before and draw a conclusion.
+
+\subsection{Observational Determinism for Concurrent Program Security}
+\subsection{Flexible Scheduler-Independent Security}
 

### 2012-01-28 paper/paper.tex
added first two paragraphs of FSI

@@
 \subsection{Observational Determinism for Concurrent Program Security}
-\subsection{Flexible Scheduler-Independent Security}
+\subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}
+This approach developed by Mantel and Sudbrock has the goal to be on the one hand
+scheduler independent and on the other not too restrictive (i.e. does not treat
+obviously secure programs as insecure).
+
+To reach this goal they model the scheduling explicit but general enough to apply
+to a wide class of common schedulers. So rather than talking about program states,
+they introduces system configurations which contain the list of threads the global
+memory and the scheduler state. The scheduler is able to store information in its
+state and can retrieve input (e.g. the number of active threads). Based on this
+information the scheduler makes decisions. To cover probabilistic schedulers the
+decision making of the scheduler and the execution of the program is probabilistic.
 

### 2012-01-28 paper/paper.tex
added citation "obversavtional determinism…"
added subsection for the conclusion
added whitespace

@@
 In this summary I will follow mainly the structure of the original
-paper\cite{Barthe07} which is as follows. After the introduction the basic
+paper \cite{Barthe07} which is as follows. After the introduction the basic
 terms and definitions for a multithreaded programs and the scheduler are
@@
 
-\subsection{Observational Determinism for Concurrent Program Security}
+\subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
 \subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}
@@
 
+\subsection{Conclusion}
+
 \bibliography{bibliography}

### 2012-01-28 paper/paper.tex
made references to typing rules universal

@@
 \subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
-by the following rules.
+by the rules in figure \ref{fig:multithreaded-execution}.
 
@@
 
-We now take a look at the rules of this type system concerning the \texttt{if} instruction.
+We now take a look at the rules of this type system concerning the \texttt{if} instruction, which are
+defined in figure \ref{fig:typing-rules-if}.
 

### 2012-01-29 paper/paper.tex
s/probabilistic/nondeterministic
added rest of FSI-security

@@
 to a wide class of common schedulers. So rather than talking about program states,
-they introduces system configurations which contain the list of threads the global
+they introduces system configurations which contain the list of threads (thread pool) the global
 memory and the scheduler state. The scheduler is able to store information in its
 state and can retrieve input (e.g. the number of active threads). Based on this
-information the scheduler makes decisions. To cover probabilistic schedulers the
-decision making of the scheduler and the execution of the program is probabilistic.
+information the scheduler makes decisions. To cover nondeterministic schedulers
+(e.g. the uniform scheduler) the decision making of the scheduler and the execution
+of the program is probabilistic.
+
+Based on this scheduler model, a schedule-specific security property $\mathcal{S}$ is
+defined on thread pools. Intuitively a thread pool is $\mathcal{S}$-secure if the
+probability of running a program with system configuration $conf$ under the scheduler $\mathcal{S}$
+results in the same global memory modulo low-equality as running the program with
+system configuration $conf'$.
+
+The novel security property introduced in this paper is defined without mentioning the
+scheduler and equal to $\mathcal{S}$-security for a \textit{robust} scheduler $\mathcal{S}$.
+Therefor this property is called \textit{flexible scheduler-independent security} (FSI-security).
+A thread pool is FSI-secure if (starting from an low-equal memory configuration) the resulting
+memories are always low-equal when executing commands that potentially modify low-variables and if
+spawned threads are also FSI-secure.
+
+Those robust schedulers are those, for which \textit{"the scheduling of low threads during a run of a FSI-secure
+thread pool remains unchanged when one removes all high threads from the thread pool"}\cite{Mantel10}.
+
+If a program is FSI-secure or not is checked statically with a security type system, which is introduced
+for a sample while-language.
 

### 2012-01-29 paper/paper.tex
fixed direct citations

@@
 
-The idea is to 'track and regulate' \cite{Zdancewic04} where information flows to
+The idea is to \textit{"track and regulate"}\cite{Zdancewic04} where information flows to
 prevent the leak of secret information. The ways on which the information flows
@@
 The notion of confidentiality that is used in most work in the area of information-flow security is \textit{noninterference}.
-This policy is defined in various ways but the essence is, that it requires 'that secret information
-[do] not affect [the] publicly observable behavior of a system'\footnote{Even wen the attacker has full access to the
+This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
+[do] not affect [the] publicly observable behavior of a system"}\footnote{Even wen the attacker has full access to the
 source code of the program.}\cite{Zdancewic04}. For most real world
@@
 The next thing introduced are the security levels. We assume the attacker
-'is' a level $k$. From this assumption we can reduce every set of security levels
+"is" a level $k$. From this assumption we can reduce every set of security levels
 w.l.o.g into $\{low, high\}$, where $low < high$, by mapping elements
@@
 these hidden threads, the scheduler will be modified to treat these threads
-in a special way. This will be done by 'hiding' these threads, therefore comes
+in a special way. This will be done by "hiding" these threads, therefore comes
 the name of them.
@@
 security level that was assigned by $\Gamma$ to the variable $x$. If this is the case the head
-of $LType$ is removed. The $\sqcup$ operator is like a logical 'or', it returns the higher of
+of $LType$ is removed. The $\sqcup$ operator is like a logical "or", it returns the higher of
 the given security values.

### 2012-01-29 paper/paper.tex
added section about observational determinism

@@
 \subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
+Zdancewic and Myers approach has to goal to provide a secure concurrent language
+that has a general and realistic support for concurrency and whose security can
+be checked statically.
+
+They introduced the language $\lambda_{SEC}^{PAR}$ which supports higher-order functions, an
+unbounded number of threads, synchronization (via join patterns), message passing and shared memory. Regardless of those realistic
+features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-lever programming language because its
+syntax and type system is too awkward, instead it is a model for studying information flow.
+
+Low-security observational determinism is used to ensure noninterference. The idea is
+that to make the naturally nondeterministic system, deterministic from the point of view
+of the attacker.
+
+This deterministic noninterference is defined as
+
+\begin{align*}
+(m \approx_\zeta m' \land m \Downarrow T \land m' \Downarrow T') \Rightarrow T \approx_\zeta T'
+\end{align*}
+
+Where $T$ and $T'$ are traces of the execution of a program. $m$ and $m'$ are initial configurations.
+Let $T(L) = [M_0(L), M_1(L), \dots]$ be the list of values at location $L$ in the trace $T$. Then
+$T \approx_\zeta T'$ holds if the values of $T(L)$ and $T'(L)$ are pairwise indistinguishable. Since
+it is allowed that one sequence is a prefix of the other, there is no need to proof termination, but
+external timing attacks are possible.
+
+This notion of observational determinism is captured in a type system for $\lambda_{SEC}^{PAR}$, thus
+the security of programs can be checked statically.
+
 \subsection{Flexible Scheduler-Independent Security\cite{Mantel10}}

### 2012-01-29 paper/paper.tex
added conclusion

@@
 \subsection{Conclusion}
+The approach of Barthe et. al. shows how to build a framework to create type-annotated programs that can
+be check statically and can be run independent of the scheduling algorithm. Thus there is no need to trust
+the compiler.
+
+In contrast to Mantel and Sudbrocks work the whole framework depends heavily on the security of the scheduler.
+It is possible to choose most scheduling algorithms, but every algorithm needs to be modified satisfy the
+properties for a secure scheduler. Additionally its likely that the interface of the scheduler needs to be
+changed/extended in nearly ever implementation. The FSI-security instead is applicable to every robust
+scheduler without changing the algorithm or the interface.
+
+The approach also lacks (currently) support for real world languages. For example support of synchronization
+would be necessary to implement the approach for Java. The approach of Zdancewic and Myers supports those
+features, but is much more restrictive. Due to the use of observational determinism obviously secure programs
+like \texttt{l := 1 || l := 0} are rejected. The approach of Barthe et. al. does not reject those programs and
+is in that sense much more permissive.
+
+In the end Barthe et. al. presented a framework with reaches many goals (like permissiveness, statically check-able
+and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
+hard to use for real programs.
 

### 2012-01-29 paper/paper.tex
used names of the paper for the next properties
made next properties \small to fit in \textwidth
added note aboute \mapsto^*
corrected spelling
use new property names

@@
 
-\todo{Use names from paper}
+{\small
 \begin{enumerate}
-\item $Dom(next) = \{i \in \mathcal{P} | se(i) = high \land \exists j \in \mathcal{P} . i \mapsto^* j \land se(j) \neq high\}$ \todo{use simple method of new paper?}
-\item $i, j \in Dom(next) \land i \mapsto j \Rightarrow next(i) = next(j)$
-\item $i \in Dom(next) \land j \not\in Dom(next) \land i \mapsto j \Rightarrow next(i) = j$
-\item $j, k \in Dom(next) \land i \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = next(k)$ \todo{is there an error?}
-\item $i, j \in Dom(next) \land k \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = k$
+\item[NePd] $Dom(next) = \{i \in \mathcal{P} | se(i) = high \land \exists j \in \mathcal{P} . i \mapsto^* j \land se(j) \neq high\}$ \todo{use simple method of new paper?}
+\item[NeP1] $i, j \in Dom(next) \land i \mapsto j \Rightarrow next(i) = next(j)$
+\item[NeP2] $i \in Dom(next) \land j \not\in Dom(next) \land i \mapsto j \Rightarrow next(i) = j$
+\item[NeP3] $j, k \in Dom(next) \land i \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = next(k)$ \todo{is there an error?}
+\item[NeP4] $i, j \in Dom(next) \land k \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = k$
 \end{enumerate}
+}
 
-Die domain of next just captures the high, but not always high program points (i.e. the ones
-that can result in a hidden thread). The second property, states that two directly successive high
-program points have the same program point in which the thread becomes visible again. The third rule
-is the counterpart of the second: If a low program point follows a high program point, this low program
+where $\mapsto^*$ is the reflexive and transitive closure of $\mapsto$.
+
+The domain of next (i.e. NePd) captures the high, but not always high program points (i.e. the ones
+that can result in a hidden thread). The property NeP1, states that two directly successive high
+program points have the same program point in which the thread becomes visible again. The property NeP2
+is the counterpart of the NeP1: If a low program point follows a high program point, this low program
 point is the result of the next function for the high program point.

### 2012-01-29 paper/paper.tex
explained NeP3 and NeP4 properties
removed todos
removed todo
fixed spelling

@@
 
-\todo{Explain intuition}
-
 \subsection{Soundness}
@@
 \begin{enumerate}
-\item[NePd] $Dom(next) = \{i \in \mathcal{P} | se(i) = high \land \exists j \in \mathcal{P} . i \mapsto^* j \land se(j) \neq high\}$ \todo{use simple method of new paper?}
+\item[NePd] $Dom(next) = \{i \in \mathcal{P} | se(i) = high \land \exists j \in \mathcal{P} . i \mapsto^* j \land se(j) \neq high\}$
 \item[NeP1] $i, j \in Dom(next) \land i \mapsto j \Rightarrow next(i) = next(j)$
 \item[NeP2] $i \in Dom(next) \land j \not\in Dom(next) \land i \mapsto j \Rightarrow next(i) = j$
-\item[NeP3] $j, k \in Dom(next) \land i \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = next(k)$ \todo{is there an error?}
+\item[NeP3] $j, k \in Dom(next) \land i \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = next(k)$
 \item[NeP4] $i, j \in Dom(next) \land k \not\in Dom(next) \land i \mapsto j \land i \mapsto k \land j \neq k \Rightarrow next(j) = k$
@@
 is the counterpart of the NeP1: If a low program point follows a high program point, this low program
-point is the result of the next function for the high program point.
-
-\todo{What is wrong with property 4 (NeP3)}
-\todo{What is wrong with property 5 (NeP4)}
+point is the result of the next function for the high program point. NeP3 and NeP4 denote that the
+next of an if respective while instruction is at least after the control dependence region.
 
@@
 
-The approach also lacks (currently) support for real world languages. For example support of synchronization
+The approach also (currently) lacks support for real world languages. For example support of synchronization
 would be necessary to implement the approach for Java. The approach of Zdancewic and Myers supports those

### 2012-01-29 paper/paper.tex
revied whole paper

@@
 a network connection). Through the timing difference of the scheduled threads it
-is likely that sensitive information is leaked and an attack may obtain all
+is likely that sensitive information is leaked and an attacker may obtain
 sensitive data.
@@
 The idea Barthe, Rezk, Russo and Sabelfeld phrased in \cite{Barthe07} was to close
-this leak (more formally called \textit{covert channel}) by annotating the byte-code
-of the program (most mobile devices use some sort of byte-code e.g. Android) with security
-labels according to the processed information, which causes the scheduler to hide
+this leak (formally called \textit{covert channel}) by annotating the byte-code
+of the program\footnote{most mobile devices use some sort of byte-code e.g. Android} with security
+labels according to the confidentiality of the information, which causes the scheduler to hide
 threads that leak information from the attacker.
@@
 a computer lead thirty years ago (c.f. \cite{Zdancewic04}) to the research
-on the security information-flows, because of the shortcomings of the standard methods.
+on the security of information-flows, because of the shortcomings of the standard methods.
 These standard methods are: Access control and encryption. Where
@@
 be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
-variable and $l$ a variable that is plublic):
+variable and $l$ a variable that is public):
 
@@
       is leaked through the control structure of the program (e.g. $\text{if}\ h = 1\ \text{then}\ l := 1\ \text{else}\ l := 0$).
-\item \textit{Probabilistic channels}: If an attacker is able to run a computation multiple times
+\item \textit{Probabilistic channels}: If an attacker is able to run a computation multiple times,
       he might be able to obtain information by looking at the probability distribution of the
       public outputs.
-\item \textit{Power channels}: If the attacker has physical access to the computer or at least figure out
+\item \textit{Power channels}: If the attacker has physical access to the computer or at least can figure out
       the power consumption, he might be able to obtain data through the changing power consumption.
@@
 \item \textit{Timing channels:} Information can be obtained through the time at which this action
-      occurs. On the one hand \textit{external} timing channels cover the action like the termination
+      occurs. On the one hand \textit{external} timing channels cover the actions like the termination
       of the program, where the attacker obtains information from the total execution time of the program.
-      On the other hand \textit{internal} timing channels are occur in multithreaded programs through the timing difference
+      On the other hand \textit{internal} timing channels occur in multithreaded programs through the timing difference
       between threads. For example the two threaded (denoted with $||$) program \[\text{if}\ h = 1\ \{\text{sleep}(100)\}\ ;\ l := 1\ ||\ \text{sleep}(50)\ ;\ l := 0\]
       executed with most schedulers will leak the information of $h$ into $l$ through the timing difference
-      if $h = 1$.
+      between $h = 1$ and $h = 0$.
 \end{itemize}
@@
 This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
-[do] not affect [the] publicly observable behavior of a system"}\footnote{Even wen the attacker has full access to the
-source code of the program.}\cite{Zdancewic04}. For most real world
+[do] not affect [the] publicly observable behavior of a system"}\cite{Zdancewic04}\footnote{Even if the attacker has full access to the
+source code of the program.}. For most real world
 applications this policy is far too strict, because it forbids useful programs like password checkers.\footnote{A
@@
 
-Mechanisms for controlling the information flow can be implemented either dynamic or static. The dynamic approach\footnote{
+Mechanisms for controlling the information flow can be implemented either dynamic or static. Dynamic approaches\footnote{
 The taint mode of the programming language Perl uses this mechanism.}
-labels the information with security labels and propagates these labels wherever the information is used. The static
-approach\footnote{This approach is implemented in Jif for Java and Flow Caml for Caml.} analyses the program code and is
-therefor far more promising, because with this approach it is possible to check all evaluation paths.
+label the information with security labels and propagate these labels wherever the information is used. Static
+approaches\footnote{This mechanism is implemented in Jif for Java and Flow Caml for Caml.} analyze the program code and are
+therefor far more promising, because with these approaches it is possible to check all evaluation paths.
 
@@
 In this summary I will follow mainly the structure of the original
-paper \cite{Barthe07} which is as follows. After the introduction the basic
-terms and definitions for a multithreaded programs and the scheduler are
+paper \cite{Barthe07} which is as follows. First the basic
+terms and definitions for multithreaded programs and the scheduler are
 laid. After that the notion of security we want to achieve is presented.
 Along with this a skeleton of a type system is described, which ensures
-that a program typable in this type system is secure w.r.t the notion of
-security. The proof that this holds is sketched in the following section.
+that a program typeable in this type system is secure.
+The proof that this holds is sketched in the following section.
 In the last section the (by now) abstract framework is instantiated with a
@@
 to create a new thread (\texttt{start \textit{pc}} where pc is the start
-instruction of the new thread).
+instruction of the new thread). The instructions without \texttt{start}
+are called $SeqIns$.
 
@@
 "is" a level $k$. From this assumption we can reduce every set of security levels
-w.l.o.g into $\{low, high\}$, where $low < high$, by mapping elements
+w.l.o.g into $\text{Level} = \{low, high\}$, where $low < high$, by mapping elements
 that are no more sensitive than $k$ to $low$ and all other elements
@@
 with respect to their current program point (\texttt{s.pc(tid)} where $tid \in \text{Thread}$),
-into \textit{low} threads iff the current program point is low. In \textit{high}
+into \textit{low} threads if the current program point is low. In \textit{high}
 threads iff the current program point is high. In \textit{always high} threads
-iff the current program point is always high and in \textit{hidden} threads, iff
+if the current program point is always high and in \textit{hidden} threads, if
 the current program point is high but not always high.
@@
 The latter two are the interesting ones. If a thread is \textit{always high} it
-can not leaks any information into low, because it never gets in touch
-with low program points. So these threads can safely be interleaved between any
+can not leak any information into low, because it never gets in touch
+with low program points. So these threads can safely be interleaved between all
 other threads by the scheduler.
@@
 At this point no concrete scheduler is defined to make the framework applicable
-for a wide class of schedulers. A scheduler is in this class iff it can be
+for a wide class of schedulers. A scheduler is in this class if it can be
 modeled as a function $pickt: \text{ConcState} \times \text{History} \rightharpoonup \text{Thread}$
@@
 \item It always picks an active thread
-\item If there is a hidden thread, always choose high or always high threads
+\item If there is a hidden thread, always choose high or always high threads\footnote{With this constraint the interleaving of always high threads is realized}
 \item Only use low information and the low part of the history to choose a
@@
 
-In constraint two the interleaving of always high threads is realized.
-
-To define what noninterference means we need a notion of excution. It is assumed
+To define what noninterference means we need a notion of execution. It is assumed
 that we have a sequential execution relation $\leadsto_{seq} \subseteq \text{SeqState} \times
-\text{SeqState}$. The one step execution for the multithreaded language $\leadsto_{seq}
+\text{SeqState}$. One step execution for the multithreaded language $\leadsto_{seq}
 \subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
@@
 
-The scheduler is allow to pick a new thread after every transition of $\leadsto_{seq}$. The
-intuition of the first rule is, that the scheduler picks a new thread $ctid$ which is at
-program point $i$. This program point maps to an sequential instruction. The exection of
-this instruction leads to a local state $\sigma$ and a global memory $\mu$. If these conditions
-are met and instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
-then the concurent transition can be made. In this transition the concurrent state is updated
-according to the results of the sequential execution and the thread id and security level of $i$
+The scheduler is allowed to pick a new thread after every transition of $\leadsto_{seq}$. The
+intuition of the first rule is, that the scheduler picks a new thread $coed$ which is at
+program point $i$. This program point maps to an sequential instruction. The execution of
+this instruction leads to a local state $\sigma$ and a global memory $\mu$. If this is the case
+and the instruction $i$ was not the last instruction (i.e. $\sigma.pc \neq \mathtt{exit}$),
+then the concurrent transition can be made. In this transition the concurrent state is updated
+according to the results of the sequential execution and the thread id furthermore security level of $i$
 is recorded in the history.
 
-The second rule covers the case in which the thread pickted by the scheduler has only the current
+The second rule covers the case in which the thread picked by the scheduler has only the current
 instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
@@
 If the current instruction at program point $i$ is the \texttt{start} instruction, a new thread
-identifier is picked w.r.t. the security level of $i$. The program counter of the current thread
+identifier is picked w.r.t. the security level of $i$, then program counter of the current thread
 is increased by one, to step to the next instruction. The resulting concurrent state includes the
@@
 
-So the state $s$ evaluates according to a history $h$ to a final global memory $\mu$ iff there is a
-sequence of concurrent executions that terminates (i.e. there are no active threads) and has the
+The intuition is that the state $s$ evaluates according to a history $h$ to a final global memory $\mu$ iff there is a
+sequence of concurrent executions that terminates (i.e. there are no active threads left) and has the
 global memory $\mu$. $\leadsto_{conc}^*$ is the reflexive and transitive closure of of $\leadsto_{conc}$.
@@
 in accordance to an indistinguishability relation $\sim_g$ on global memories.
-Barthe et.al. state that it is not necessary, for the purpose of the paper,
-to specify the definition of this relation. But to get a feel for this
-relation, you can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
+Barthe et. al. state that it is not necessary -- for the purpose of the paper --
+to specify the definition of this relation. But to get a feeling for this
+relation, one can define it as: $\mu \sim_g \mu' \Leftrightarrow \mu|_{low} =
 \mu'|_{low}$ where $\mu|_{low}$ projects out all high elements. Based on that
-a program P is noninterfering if for all gobal memories $\mu_1, \mu_2, \mu_1'$
+a program P is non-interfering if for all global memories $\mu_1, \mu_2, \mu_1'$
 and $\mu_2'$ it holds that:
@@
 \begin{align*}
-\mu_1 \sim_g \mu_2 \land P,\mu_1 \Downarrow_{conc} \mu_1' \land P,\mu_2 \Downarrow_{conc} \mu_2' \Rightarrow \mu_1' \sim_g \mu_2'
+\mu_1 \sim_g \mu_2 \land (P,\mu_1 \Downarrow_{conc} \mu_1') \land (\text{P},\mu_2 \Downarrow_{conc} \mu_2') \Rightarrow \mu_1' \sim_g \mu_2'
 \end{align*}
@@
 \label{sec:typesystem}
-The type system is the core part of the framework in the effect that it
+The type system is the core part of the framework in the sense that it
 enforces the previously defined noninterference property. Thus every
-program which is typable, is noninterfering.
+program which is typeable, is non-interfering.
 
@@
 
-The intuition of the typing judgment is that if we execute the instruction at program
-point $i$ w.r.t. the security environment $se$ and the current type is $\textsc{s}$. Then the
+The intuition of the typing judgment is, that if we execute the instruction at program
+point $i$ w.r.t. the security environment $se$ and the current type is $\textsc{s}$, then the
 type after the execution is $\textsc{t}$.
@@
 and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point
-(including the ones of spawned threads) to $\textsc{t}_{init}$ and for every program point $j$ which is an
-successor of $i$ there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
+\footnote{including the ones of spawned threads} to $\textsc{t}_{init}$ and that for every program point $j$, which is an
+successor of $i$, there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
 by $\mathcal{S}(j)$ and $se, i \vdash \mathcal{S}(i) \Rightarrow \textsc{s}$ holds.
@@
 \begin{theorem}
-If the scheduler is secure and $se, \mathcal{S} \vdash P$, then P is noninterfering.
+If the scheduler is secure and $se, \mathcal{S} \vdash \text{P}$, then $\text{P}$ is non-interfering.
 \end{theorem}
@@
 point is the result of the next function for the high program point. NeP3 and NeP4 denote that the
-next of an if respective while instruction is at least after the control dependence region.
+next of an outermost if respective while instruction is at least after the control dependence region.
 
@@
 operational semantics are standard and not explicitly necessary for the following
-instantiation, therefor they are omited.
+instantiation, therefor they are omitted.
 
@@
 $\text{LType} = \text{Stack}(\text{Level})$ and $\textsc{t}_{init}$ with the empty stack. The typing rules defined in
-figure~\ref{fig:multithreaded-typing-rules} need to be extended with rules for the concrete instructions
-of the assembly language. In this summary I only explain two of them, the rest follows
+figure~\ref{fig:multithreaded-typing-rules} are extended with rules for the concrete instructions
+of the assembly language in figure~\ref{fig:transfer-rules}. In this summary I only explain two of them, the rest follows
 in a similar manner.
@@
 $x$ a security level. If we now want to \texttt{store} the top of the stack into the variable
-$x$ one premise is that the security level of the program point ($se(i)$) \textit{and} the
+$x$, one premise is that the security level of the program point $se(i)$ \textit{and} the
 security level of the value on top of the operand stack is \textit{lower or equal} to the
@@
 Now we can define $tregion$ according to a compilation function $\mathcal{C}$. I won't define
-the compilation function here, because it is not that essential how it exactly looks like.
+the compilation function here, because it is not essential how it exactly looks like.
 
@@
 In this section I will first give a short overview of two other approaches and then
-I will compare them with the approach presented before and draw a conclusion.
+compare them with the approach presented before and draw a conclusion.
 
@@
 features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-lever programming language because its
-syntax and type system is too awkward, instead it is a model for studying information flow.
+syntax and type system is too awkward. Instead it is only used a model for studying information flow.
 
-Low-security observational determinism is used to ensure noninterference. The idea is
-that to make the naturally nondeterministic system, deterministic from the point of view
+Low-security observational determinism is used to enforce noninterference. The idea is
+to make the naturally nondeterministic system, deterministic from the point of view
 of the attacker.
@@
 This approach developed by Mantel and Sudbrock has the goal to be on the one hand
-scheduler independent and on the other not too restrictive (i.e. does not treat
-obviously secure programs as insecure).
+scheduler independent and on the other not too restrictive.
 
@@
 to a wide class of common schedulers. So rather than talking about program states,
-they introduces system configurations which contain the list of threads (thread pool) the global
+they introduce system configurations which contains a list of threads (thread pool) the global
 memory and the scheduler state. The scheduler is able to store information in its
@@
 probability of running a program with system configuration $conf$ under the scheduler $\mathcal{S}$
-results in the same global memory modulo low-equality as running the program with
-system configuration $conf'$.
+resulting in a global memory $m$ is the same as running it with configuration $conf'$ resulting in $m'$,
+and satisfying that $m =_L m'$.
 
@@
 and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
-hard to use for real programs.
+hard to use for real programs. In addition they did not how to implement the type-annotations in a real
+language.
 

### 2012-01-30 paper/paper.tex
moved some sections to new pages
fixed spelling
corrected errors found by friends

@@
 this leak (formally called \textit{covert channel}) by annotating the byte-code
-of the program\footnote{most mobile devices use some sort of byte-code e.g. Android} with security
+of the program\footnote{Most mobile devices use some sort of byte-code e.g. Android.} with security
 labels according to the confidentiality of the information, which causes the scheduler to hide
@@
 prevent the leak of secret information. The ways on which the information flows
-through the system are called channels. Channels that are not indented to transport
-information are called covered channels. These covered channels can
+through the system are called \textit{channels}. Channels that are not intended to transport
+information are called \textit{covered channels}. These covered channels can
 be classified into the following categories \cite{Sabelfeld03} (in the following examples $h$ is a secret
@@
 
-The notion of confidentiality that is used in most work in the area of information-flow security is \textit{noninterference}.
+The notion of confidentiality that is used in most work in the area of in\-for\-ma\-tion-flow security is \textit{noninterference}.
 This policy is defined in various ways but the essence is, that it requires \textit{"that secret information
@@
 program points $\mathcal{P}$ with a distinguished entry ($1$) and exit
-(\texttt{exit}) point and a function $P$ that maps program points to
+point (\texttt{exit}) and a function $P$ that maps program points to
 instructions.
@@
 into \textit{low} threads if the current program point is low. In \textit{high}
-threads iff the current program point is high. In \textit{always high} threads
+threads if the current program point is high. In \textit{always high} threads
 if the current program point is always high and in \textit{hidden} threads, if
@@
 
-The latter two are the interesting ones. If a thread is \textit{always high} it
+The last two are the interesting ones. If a thread is \textit{always high} it
 can not leak any information into low, because it never gets in touch
@@
 The \textit{hidden} threads are the ones we have to care about. These contain
-obviously high information in the current program point, but have afterwards
+obviously high information in the current program point, but have subseqent
 instructions that deal with low information. Since the attacker can watch the
@@
 \item It always picks an active thread
-\item If there is a hidden thread, always choose high or always high threads\footnote{With this constraint the interleaving of always high threads is realized}
+\item If there is a hidden thread, always choose high or always high threads\footnote{With this constraint the interleaving of always high threads is realized.}
 \item Only use low information and the low part of the history to choose a
@@
 \begin{align*}
-\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] = \mathtt{start} pc \\
+\inferrule{pick(s, h) = ctid \\ s.pc(ctid) = i \\ P[i] = \mathtt{start}\ pc \\
 fresht_{se(i)}(s) = ntid \\ s(ctid).[pc := i + 1] = \sigma'}
@@
 then the concurrent transition can be made. In this transition the concurrent state is updated
-according to the results of the sequential execution and the thread id furthermore security level of $i$
+according to the results of the sequential execution and the thread identifier, furthermore security level of $i$
 is recorded in the history.
@@
 instruction before terminating (i.e. $\sigma.pc = \mathtt{exit}$). In the concurrent transition
-the current thread id is removed from the concurrent state and everything else is like in the
+the current thread identifier is removed from the concurrent state and everything else is like in the
 first rule.
@@
 updated state of the current thread and the initial state for the new thread. The global memory
-is not modified and the history is extended with the current thread id and the security level of $i$.
+is not modified and the history is extended with the current thread identifier and the security level of $i$.
 
@@
 
-This type system is extended by the following rules, to support multithreading:
+This type system is extended by the following rules to support multithreading:
 
@@
 that maps a local type to every program point
-and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point
-\footnote{including the ones of spawned threads} to $\textsc{t}_{init}$ and that for every program point $j$, which is an
+and $se$ a security environment, iff $\mathcal{S}$ maps every initial program point\footnote{
+Including the ones of spawned threads.} to $\textsc{t}_{init}$ and that for every program point $j$, which is an
 successor of $i$, there is a type $s \in \text{LType}$, such that $\textsc{s}$ is lower bounded
@@
 
+\newpage
 \subsection{Instantiation}
@@
 To demonstrate how the framework can be used, it was instantiated with a simple
-assembly language, which is given by the follow grammar:
+assembly language, which is given by the following grammar:
 
@@
 
-According to this labels we can define the dependence regions for the source language:
+According to this labels we can define the control dependence regions for the source language:
 
@@
 \begin{definition}
-For every branching point $c$ in the source program such that $\vdash_\circ [c]_\bullet^n$, the next function
+For every branching point $c$ in the source program such that $\vdash_\circ~[c]_\bullet^n$, the next function
 is defined as $\forall k \in tregion(n) .\ next(k) = jun(n)$.
@@
 
+\newpage
 \section{Comparison with other approaches/further work}
@@
 \subsection{Observational Determinism for Concurrent Program Security\cite{Zdancewic03}}
-Zdancewic and Myers approach has to goal to provide a secure concurrent language
+Zdancewic and Myers approach has the goal to provide a secure concurrent language
 that has a general and realistic support for concurrency and whose security can
@@
 and requires no intervention from the programmer) but needs to modify the scheduler interface, which makes it
-hard to use for real programs. In addition they did not how to implement the type-annotations in a real
+hard to use for real programs. In addition they did not showed how to implement the type-annotations in a real
 language.
 
+\newpage
 \bibliography{bibliography}

### 2012-02-01 paper/paper.tex
use package hyperref instead of url
some more corrections

@@
 \usepackage{nameref}
-\usepackage{url}
+\usepackage{hyperref}
 
@@
 \begin{definition}
-The junction points are computed by a function $jun: \mathcal{P} \rightarrow \mathcal{P}$. It is defined
+The junction points are computed by a function $jun: \mathcal{P} \rightarrow \mathcal{P}$. This function is defined
 on all junction points $[c]^n$ in the source program and its domain is $jun(n) = max \{i\ |\ i \in tregion(n)\} + 1$.
@@
 The first rule \textsc{L-Cond} from Figure \ref{fig:typing-rules-if} says that if we branch on a low
-guard, everything depends on the security level of $c$ and $c'$. The rule \textsc{H-Cond} covers the
+guard, then everything depends on the security level of $c$ and $c'$. The rule \textsc{H-Cond} covers the
 case where we have a high guard, in this case the control dependence region has to be marked
@@
 
-Where $T$ and $T'$ are traces of the execution of a program. $m$ and $m'$ are initial configurations.
-Let $T(L) = [M_0(L), M_1(L), \dots]$ be the list of values at location $L$ in the trace $T$. Then
+where $T$ and $T'$ are traces of the execution of a program. $m$ and $m'$ are initial configurations.
+Let $T(L) = [M_0(L), M_1(L), \dots]$ be the list of values at location $L$ in the trace $T$.
 $T \approx_\zeta T'$ holds if the values of $T(L)$ and $T'(L)$ are pairwise indistinguishable. Since

### 2012-02-01 paper/paper.tex
removed citation from title
removed remark

@@
 \title{{\normalsize Seminar: Formal Specification} \\[1ex]
-  Security of Multithreaded Programs by Compilation\cite{Barthe07}}
+  Security of Multithreaded Programs by Compilation}
 \author{Pascal Wittmann, Advisor: Artem Starostin}
@@
 The proof that this holds is sketched in the following section.
-In the last section the (by now) abstract framework is instantiated with a
+In the last section the abstract framework is instantiated with a
 concrete example.

### 2012-02-01 paper/paper.tex
corrected sentence
extended motivation

@@
 labels according to the confidentiality of the information, which causes the scheduler to hide
-threads that leak information from the attacker.
+threads that leak information from the attacker. Their formal model allows to
+formulate this idea as a framework, which can be instantiated with many programming languages,
+schedulers and types of byte-code.
 
@@
 The junction points are computed by a function $jun: \mathcal{P} \rightarrow \mathcal{P}$. This function is defined
-on all junction points $[c]^n$ in the source program and its domain is $jun(n) = max \{i\ |\ i \in tregion(n)\} + 1$.
+on all junction points $[c]^n$ in the source program as $jun(n) = max \{i\ |\ i \in tregion(n)\} + 1$.
 \end{definition}

### 2012-02-01 paper/paper.tex
removed unecessary gabs before formulas
added \ref to figure 2
corrected footnote in motivation

@@
 this leak (formally called \textit{covert channel}) by annotating the byte-code
-of the program\footnote{Most mobile devices use some sort of byte-code e.g. Android.} with security
+of the program\footnote{Most operating systems for mobile devices use some sort of byte-code (e.g. Android) for application software.} with security
 labels according to the confidentiality of the information, which causes the scheduler to hide
@@
 \text{GMemory})$ is defined by
-
 \begin{align*}
@@
 and $\mu_2'$ it holds that:
-
 \begin{align*}
@@
 
-This type system is extended by the following rules to support multithreading:
+This type system is extended by the rules in figure~\ref{fig:multithreaded-typing-rules} to support multithreading.
 
@@
 is defined as follows:
-
 \begin{align*}
@@
 source language at the regions at risk with natural numbers:
-
 \begin{align*}

### 2012-02-03 paper/paper.tex
removed duplicate 'of'
corrected subscript of relation
fixed typo

@@
 that we have a sequential execution relation $\leadsto_{seq} \subseteq \text{SeqState} \times
-\text{SeqState}$. One step execution for the multithreaded language $\leadsto_{seq}
+\text{SeqState}$. One step execution for the multithreaded language $\leadsto_{conc}
 \subseteq (\text{ConcState} \times \text{History}) \times (\text{ConcState} \times \text{History})$ is defined
@@
 sequence of concurrent executions that terminates (i.e. there are no active threads left) and has the
-global memory $\mu$. $\leadsto_{conc}^*$ is the reflexive and transitive closure of of $\leadsto_{conc}$.
+global memory $\mu$. $\leadsto_{conc}^*$ is the reflexive and transitive closure of $\leadsto_{conc}$.
 
@@
 unbounded number of threads, synchronization (via join patterns), message passing and shared memory. Regardless of those realistic
-features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-lever programming language because its
+features $\lambda_{SEC}^{PAR}$ is not intended to server as a user-level programming language because its
 syntax and type system is too awkward. Instead it is only used a model for studying information flow.
